### 2020-03-22 ProjectProposal.Rmd
added project proposal

---
title: "Inference of seasonal forcings on reproduction numbers during epidemics"
author: "Lu Lu, Ziyao Tian, Fengyang Wang, Yi Xiao Zhang"
output:
  pdf_document: default
  html_notebook: default
---

## Background

Exponential growth of the number of symptomatic cases is usually observed in the initial stages of infectious disease epidemics. Typically, the rate of this growth is decomposed into two quantities:

- The serial interval: time between onset of symptoms of two adjacent individuals in a transmission chain
- The reproductive number, $R$: the expected number of new infections a single symptomatic individual will cause

The estimation of the latter quantity is important for policy, because infectious disease outbreaks will switch from an exponential growth regime to an exponential decay regime when the reproductive number falls below 1. The goal of mitigation or suppression policies is therefore to effect change on the reproductive number $R$.

Many infectious diseases are believed to be seasonal, including most respiratory viruses, like the SARS-CoV-2 virus implicated in the 2019–2020 coronavirus pandemic. Wang et al. (2020) have analysed how the weather of locations within China correlates with local point estimates of the reproductive number $R$. However, because their analysis is frequentist in nature, their estimates of the seasonal forcing do not come with any estimate of uncertainty.

## Bayesian Framework for Estimation of Time-Varying R

Thompson et al. (2019) outline a Bayesian framework for estimation of $R(t)$ using symptomatic case counts, decomposed into local and imported cases. The EpiEstim R package implements ideas from this paper. A simplified summary of their approach, ignoring the imported cases:

- Estimation of (discrete) distribution of serial interval: Our project does not extend this portion, which estimates the quantities ws that denote the probability density of interval length s under a posterior distribution for the serial interval.
- Estimation of time-varying $R$: Data is collected containing the number of symptomatic cases observed at time $t$: $I(t)$. The observed number of cases at each timestep is assumed to be drawn from a Poisson distribution:

$$ I(t) \space | \space  I(1), \cdots ,I(t-1),w,R(t) \sim \text{Poisson} \left(R(t) \sum_{s=1}^t I(t-s)w_s \right) $$

Then Bayesian inference can be used to compute the posterior distribution of $R(t)$ conditional on the data and the serial interval distribution. (In the Thompson et al. paper, a gamma-distributed prior for $R(t)$ [the conjugate prior for a Poisson likelihood] and a convenient serial interval distribution was chosen so that this posterior is proportional to an analytic expression.)

## Extension of Thompson et al. for $R$ subject to forcings

Our proposal is to extend the method of Thompson et al. to multiple locations, each with its own incidence time series $I_1(t),\cdots,I_n(t)$ such that

$$ R_i(t) = \beta^T x_i(t) + \varepsilon_i(t) $$

where $\beta$ is a parameter describing the locally linear effect of each seasonal forcing, $x_i(t)$ is a vector of observed seasonal factors at time $t$ and location $i$, and $\varepsilon_i(t)$ denotes the component of $R_i(t)$ not linearly predicted by seasonal conditions at location $i$.  and the serial interval distribution $(w_s)_{s \geq 1}$ are assumed to be characteristics of the disease independent of location. The observed number of cases at each timestep is

$$ I_i(t) \space | \space I_i(1), \cdots ,I_i(t-1),w,x_i(t), \varepsilon_i(t) \sim \text{Poisson} \left( (\beta^Tx_i(t)+ \varepsilon_i(t)) \sum_{s=1}^t I(t-s) w_s \right) $$

Here the parameter of interest is $\beta$ and the $\varepsilon_i(t)$ are nuisance variables. We propose to use Markov chain Monte Carlo to compute a joint distribution over all the parameters and marginalize over the $\varepsilon_i(t)$.
Because of concerns over confounding factors that may affect R and are correlated with temperature and humidity, we plan to add several other features to $x(t)$ to help control for these factors. We also need to restrict the analysis to the pre-lockdown regime in each province, because the lockdown is likely to change the magnitude of seasonal forcings.

- Time (days)
- GDP per capita
- Population density

## Dataset

- The Italian government dataset at https://github.com/pcm-dpc/COVID-19 contains the disease incidence for each province. This data contains date of confirmation instead of date of onset of symptoms, and so some procedure may be necessary to correct for this. Abbott et al. (2020) outlines a technique to correct for reporting delays using an exponential distribution of reporting delay fit to line list data; we may use the same distribution.
- We also will need to find temperature and humidity data. We have located several possible sources and are evaluating them.

## Directions for Future Research

- Looking at the climate and infection data in American and Canadian cities as the situation develops further. In particular, compare this with the results in Italy, and observe if the linear assumption between seasonal factors and $R$ holds true across different climate regions. If not, try fitting while assuming a nonlinear relationship between $R$ and seasonal factors.
- Look into statistical simulation models which consider intercity travel and geographic proximity.

## Citations

- (PRE-PRINT) Abbott, S., Hellewell, J., Munday, J. D., Chun, J. Y., Thompson, R. N., Bosse, N. I., Chan, Y. D., Russell, T. W., Jarvis, C. I., ... & Funk, S. (2020). Temporal variation in transmission during the COVID-19 outbreak. CMMID. Available at CMMID: [https://cmmid.github.io/topics/covid19/current-patterns-transmission/global-time-varying-transmission.html](https://cmmid.github.io/topics/covid19/current-patterns-transmission/global-time-varying-transmission.html)

- Thompson, R. N., Stockwin, J. E., van Gaalen, R. D., Polonsky, J. A., Kamvar, Z. N., Demarsh, P. A., ... & Lessler, J. (2019). Improved inference of time-varying reproduction numbers during infectious disease outbreaks. Epidemics, 29, 100356.

- Wang, J., Tang, K., Feng, K., & Lv, W. (2020). High Temperature and High Humidity Reduce the Transmission of COVID-19. Available at SSRN: [https://ssrn.com/abstract=3551767](https://ssrn.com/abstract=3551767) or [http://dx.doi.org/10.2139/ssrn.3551767](http://dx.doi.org/10.2139/ssrn.3551767)

### 2020-03-22 ProjectProposal.Rmd
Update source for weather

@@
 
-## Bayesian Framework for Estimation of Time-Varying R
+## Bayesian Framework for Estimation of Time-Varying $R$
 
@@
 We propose to use Markov chain Monte Carlo to compute a joint distribution over all the parameters and marginalize over the $\varepsilon_i(t)$.
-Because of concerns over confounding factors that may affect R and are correlated with temperature and humidity, we plan to add several other features to $x(t)$ to help control for these factors.
+Because of concerns over confounding factors that may affect $R$ and are correlated with temperature and humidity, we plan to add several other features to $x(t)$ to help control for these factors.
 We also need to restrict the analysis to the pre-lockdown regime in each province, because the lockdown is likely to change the magnitude of seasonal forcings.
@@
 (2020) outlines a technique to correct for reporting delays using an exponential distribution of reporting delay fit to line list data; we may use the same distribution.
-- We also will need to find temperature and humidity data. We have located several possible sources and are evaluating them.
+- The `worldmet` R package enables access to temperature and humidity data, sourced from NOAA.
 

### 2020-04-15 DataPipeline.R
scripts that fetch csvs

@@ -1,0 +1,60 @@
+library(lubridate)
+
+#' Fetches a raw file from url.
+#'
+fetch_file <- function(url, filename) {
+  tryCatch({
+    code <- download.file(url, filename)
+    if (code != 0) {
+      stop("Error downloading file")
+    }
+  },
+  error = function(e) {
+    stop(sprintf("Error downloading file '%s': %s",
+                 url, e$message))
+  })
+}
+
+#' Fetches the latest COVID-19 cases data from ECDC.
+#'
+fetch_latest_ecdc <- function() {
+  url <- "https://opendata.ecdc.europa.eu/covid19/casedistribution/csv"
+  file_name <- "data/ecdc-COVID-19-up-to-date.csv"
+  fetch_file(url, filename)
+  
+  #' Quick cleanup on the column titles.
+  #'
+  d <- read.csv(filename, stringsAsFactors = FALSE)
+  d$t <- lubridate::decimal_date(as.Date(d$dateRep, format = "%d/%m/%Y"))
+  d <- d[order(d$'countriesAndTerritories', d$t, decreasing = FALSE), ]
+  names(d)[names(d) == "countriesAndTerritories"] <- "Countries.and.territories"
+  names(d)[names(d) == "deaths"] <- "Deaths"
+  names(d)[names(d) == "cases"] <- "Cases"
+  names(d)[names(d) == "dateRep"] <- "DateRep"
+  saveRDS(d, "data/ecdc-COVID-19-up-to-date.rds")
+}
+
+#' Fetches latest data from John Hopkins University CSSE.
+#'
+fetch_latest_csse <- function() {
+  url <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv"
+  filename <- "data/time_series_covid19_confirmed_global.csv"
+  fetch_file(url, filename)
+  
+  url <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv"
+  filename <- "data/time_series_covid19_deaths_global.csv"
+  fetch_file(url, filename)
+  
+  url <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_recovered_global.csv"
[…]

### 2020-04-15 ThompsonModel.R

@@ -1,0 +1,16 @@
+#' Follows the Framework of the Thompson et al paper.
+#' Using the EpiEstim library to estimate
+#'
+
+library(EpiEstim)
+
+#' Using EpiEstim.
+#'
+#'
+#'
+
+confirmed_df <- read.csv("data/time_series_covid19_confirmed_global.csv")
+deaths_df <- read.csv("data/time_series_covid19_deaths_global.csv")
+recovered_df <- read.csv("data/time_series_covid19_recovered_global.csv")
+
+confirmed_df

### 2020-04-15 DataPipeline.R
fixed italian link

@@ -33,6 +33,7 @@ fetch_latest_ecdc <- function() {
   names(d)[names(d) == "dateRep"] <- "DateRep"
   saveRDS(d, "data/ecdc-COVID-19-up-to-date.rds")
 }
+fetch_latest_ecdc()
 
 #' Fetches latest data from John Hopkins University CSSE.
 #'
@@ -49,12 +50,13 @@ fetch_latest_csse <- function() {
   filename <- "data/time_series_covid19_recovered_global.csv"
   fetch_file(url, filename)
 }
+fetch_latest_csse()
 
-
+#' Fetch Italian provinces data.
+#'
 fetch_latest_dpc <- function() {
-  #' Fetch Italian provinces data.
-  #'
-  url <- "https://raw.githubusercontent.com/pcm-dpc/COVID-19/master/dati-province/dpc-covid19-ita-province-latest.csv"
-  filename <- "data/dpc-covid19-ita-province-latest.csv"
+  url <- "https://raw.githubusercontent.com/pcm-dpc/COVID-19/master/dati-province/dpc-covid19-ita-province.csv"
+  filename <- "data/dpc-covid19-ita-province.csv"
   fetch_file(url, filename)
 }
+fetch_latest_dpc()

### 2020-04-15 ThompsonModel.R

@@ -13,4 +13,6 @@ confirmed_df <- read.csv("data/time_series_covid19_confirmed_global.csv")
 deaths_df <- read.csv("data/time_series_covid19_deaths_global.csv")
 recovered_df <- read.csv("data/time_series_covid19_recovered_global.csv")
 
-confirmed_df
+dpc_df <- read.csv("data/dpc-covid19-ita-province-latest.csv")
+
+dpc_df

### 2020-04-15 R/ForcingDataAugmentation.R
add pipeline for demographic data

@@ -1,0 +1,86 @@
+library(restatapi)
+library(dplyr)
+library(worldmet)  # TODO : not done yet
+
+# Obtain mapping of health ministry province names to eurostat codes
+
+getprovincelist <- function() {
+  # The island of Sardinia was reorganized between collection of the demographic data and
+  # the health ministry's collection of COVID-19 data, so demographic data does not map
+  # cleanly. Because Sardinia is an island and the situation may be unique, we have just
+  # excluded it from the analysis.
+  sardiniaprefix <- "ITG2"
+
+  dsd <- get_eurostat_dsd("nama_10r_3gdp")
+  itprovinces <- select(
+    filter(
+      filter(dsd, concept == "GEO"),
+      startsWith(code, "IT") &
+        nchar(code) == 5 &
+        code != "ITZZZ" &
+        !startsWith(code, sardiniaprefix)
+    ),
+    c("code", "name")
+  )
+  # 3 provinces have names which do not match the ones the health ministry uses;
+  # so we rename them to match
+  statname <- c("Valle d'Aosta/Vallée d'Aoste", "Bolzano-Bozen", "Massa-Carrara")
+  healthname <- c("Aosta", "Bolzano", "Massa Carrara")
+  itprovinces <- left_join(
+    itprovinces,
+    data.frame(statname, healthname, stringsAsFactors = FALSE),
+    by = c("name" = "statname")
+  )
+  itprovinces$name <- coalesce(itprovinces$healthname, itprovinces$name)
+  select(itprovinces, c("code", "name"))
+}
+
+# Fetch GDP per capita data and population density data for Italian provinces
+# Source: Eurostat
+getdemodata <- function() {
+  itprovinces <- getprovincelist()
+
+  gdppercapita <- filter(get_eurostat_data(
+    "nama_10r_3gdp",
+    date_filter="2017",
+    filters=itprovinces$code
+  ), unit == "EUR_HAB")
+  gdppercapita <- select(gdppercapita, c("geo", "values"))
+  gdppercapita$geo <- as.character(gdppercapita$geo)
+
+  density <- get_eurostat_data(
+    "demo_r_d3dens",
+    date_filter="2017",
+    filters=itprovinces$code
+  )
+  density <- droplevels(select(density, c("geo", "values")))
+  density$geo <- as.character(density$geo)
+
+  df <- rename(
+    full_join(gdppercapita, density, by = 'geo'),
[…]

### 2020-04-15 ThompsonModel.R
very basic use of epiestim

@@ -1,18 +1,41 @@
 #' Follows the Framework of the Thompson et al paper.
-#' Using the EpiEstim library to estimate
 #'
 
 library(EpiEstim)
+library(ggplot2)
 
-#' Using EpiEstim.
+#' Let's use the Italian Provincial datasets.
+#'
+dpc_df <- read.csv("data/dpc-covid19-ita-province.csv")
+region_names <- unique(dpc_df$denominazione_provincia)
+
+
+#' Let's just try Rome first.
+#'
+region <- "Roma"
+region_df <- subset(dpc_df, denominazione_provincia == region, select = c("data", "totale_casi"))
+names(region_df)[1] <- "date"
+names(region_df)[2] <- "cases"
+
+#' Add a "new cases" column for per-day data. This might already be true.
+region_df <- within(region_df, new_cases <- c(0, diff(cases)))
+region_df
+
+#' Analysis Step 1: Estimate the serial interval
+#'
+#' Let's just put these placeholders in here.
+
+est_mean_si = 2.5
+est_std_si = 1
+
+#' Analysis Step 2: Estimate the reproduction rate.
 #'
 #'
-#'
+res_parametric_si <- estimate_R(region_df$new_cases,
+      method="parametric_si",
+      config = make_config(list(
+        mean_si = est_mean_si, 
+        std_si = est_std_si)))
 
-confirmed_df <- read.csv("data/time_series_covid19_confirmed_global.csv")
-deaths_df <- read.csv("data/time_series_covid19_deaths_global.csv")
-recovered_df <- read.csv("data/time_series_covid19_recovered_global.csv")
+plot(res_parametric_si, legend = FALSE)
 
-dpc_df <- read.csv("data/dpc-covid19-ita-province-latest.csv")
-
-dpc_df

### 2020-04-15 R/ForcingDataAugmentation.R
Add weather code

@@ -1,6 +1,7 @@
 library(restatapi)
 library(dplyr)
-library(worldmet)  # TODO : not done yet
+library(worldmet)
+library(memoise)
 
 # Obtain mapping of health ministry province names to eurostat codes
 
@@ -65,9 +66,9 @@ getdemodata <- function() {
   full_join(df, itprovinces, by = c("geo" = "code"))
 }
 
-# Augment Italy DPC data with weather (TODO) and demographics
+# Augment Italy DPC data with demographics
 # Also remove Sardinia provinces
-augmentDPC <- function(dpc, demodata) {
+augmentDPCdemo <- function(dpc, demodata) {
   dpc <- filter(dpc, lat != 0)  # not regional data
   dpc <- filter(dpc, denominazione_regione != "Sardegna")
 
@@ -76,11 +77,43 @@ augmentDPC <- function(dpc, demodata) {
   df
 }
 
+closestweatherstation <- function(lat, long) {
+  getMeta(lat = lat, lon = long, n = 1)$code[1]
+}
+closestweatherstationM <- memoise(closestweatherstation)
+
+importNOAAM <- memoise(importNOAA)
+
+# Augment a single group of data points with weather, using just one station code.
+augmentsingleweather <- function(group, station) {
+  print(station$station[1])
+  noaa <- importNOAAM(code = station$station[1], year = 2020)
+  print(group)
+  print(noaa)
+  result <- left_join(group, noaa, by = "date")
+  print(result)
+  select(result, -c("station"))
+}
+
+# Augment a table with date, lat, and long columns with weather data collected closest to the given time.
+augmentDPCweather <- function(dpc) {
+  # 1. Find the appropriate station code for each latitude and longitude present in data.
+  dpc <- mutate(dpc, station = mapply(closestweatherstationM, lat = lat, long = long))
+  dpc <- mutate(dpc, date = as.POSIXct(data, format="%Y-%m-%dT%H:%M:%S", tz = "Europe/Rome"))
+  dpc <- select(dpc, -c("data"))
+  group_by(dpc, station) %>% group_modify(augmentsingleweather) %>% ungroup
+}
+
 
 # test code
 demodata <- getdemodata()
 dpc <- read.csv("data/dpc-covid19-ita-province.csv")
-df <- augmentDPC(dpc, demodata)
+df <- augmentDPCdemo(dpc, demodata)
+df <- augmentDPCweather(df)
[…]

### 2020-04-15 R/ForcingDataAugmentation.R
cummax totale_casi

@@ -73,7 +73,7 @@ augmentDPCdemo <- function(dpc, demodata) {
   dpc <- filter(dpc, denominazione_regione != "Sardegna")
 
   df <- left_join(dpc, demodata, by = c("denominazione_provincia" = "name"))
-  df <- select(df, c("data", "denominazione_provincia", "lat", "long", "gdppercapita", "density"))
+  df <- select(df, c("data", "denominazione_provincia", "lat", "long", "gdppercapita", "density", "totale_casi"))
   df
 }
 
@@ -104,10 +104,18 @@ augmentDPCweather <- function(dpc) {
   group_by(dpc, station) %>% group_modify(augmentsingleweather) %>% ungroup
 }
 
+# Repair total case data to be monotonically increasing, through taking a rolling maximum.
+repairtotalcases <- function(dpc) {
+  (dpc
+   %>% group_by(denominazione_provincia)
+   %>% group_modify(~ arrange(.x, by = data) %>% mutate(totale_casi = cummax(totale_casi)))
+   %>% ungroup)
+}
 
 # test code
 demodata <- getdemodata()
 dpc <- read.csv("data/dpc-covid19-ita-province.csv")
+df <- repairtotalcases(dpc)
 df <- augmentDPCdemo(dpc, demodata)
 df <- augmentDPCweather(df)
 write.csv(df, "data/dpc-augmented.csv")

### 2020-04-15 R/DateOfInfectionInference.R
add stan initial draft

@@ -1,0 +1,6 @@
+library(EpiNow)
+
+#' Infer date of infection back in time from date of confirmation
+#' based on the fitted distribution.
+
+# TODO

### 2020-04-16 R/ClimateForcingModel.R
stan code

@@ -1,0 +1,55 @@
+library(rstan)
+library(dplyr)
+library(abind)
+
+# compile stan model
+cf_mod <- stan_model("ClimateForcingModel.stan")
+
+save(cf_mod, file="data/cf_mod.rds")
+
+df <- read.csv("data/dpc-augmented.csv", stringsAsFactors = FALSE)
+# TODO: this is probably duplicated work from somewhere
+df <- mutate(
+  df,
+  infections = coalesce(pmax(0, totale_casi - lag(totale_casi)), 0)
+)
+# TODO: this is a hack, do this a better way...
+# Drop the last date as it is NA for all locations.
+# Then drop the locations with some NA in data.
+df <- filter(df, date != max(as.character(date)))
+df <- (
+  df
+  %>% group_by(denominazione_provincia)
+  %>% filter(!any(is.na(air_temp)) & !any(is.na(RH)))
+  %>% ungroup
+)
+
+X <- abind(split(
+  select(df, c('date', 'air_temp', 'RH', 'gdppercapita', 'density')),
+  df$denominazione_provincia
+), along=3)
+
+I <- abind(split(df$infections, df$denominazione_provincia), along=2)
+
+# TODO: use a correct discretized serial interval distribution
+w <- rep(0, nrow(I))
+w[1] <- 0.1
+w[2] <- 0.2
+w[3] <- 0.3
+w[4] <- 0.2
+w[5] <- 0.1
+w[6] <- 0.1
+
+# stan data
+cf_data <- list(
+  P = ncol(X),
+  L = ncol(I),
+  T = nrow(I),
+  w = w,
+  i = I,
+  x = X
+)
+
+# fit stan model
+cf_fit <- sampling(cf_mod, data = cf_data, iter = 400,
+                   verbose = TRUE, chains = 4)

### 2020-04-16 R/ForcingDataAugmentation.R

@@ -116,8 +116,9 @@ repairtotalcases <- function(dpc) {
 demodata <- getdemodata()
 dpc <- read.csv("data/dpc-covid19-ita-province.csv")
 df <- repairtotalcases(dpc)
-df <- augmentDPCdemo(dpc, demodata)
+df <- augmentDPCdemo(df, demodata)
 df <- augmentDPCweather(df)
+write.csv(demodata, "data/demodata.csv")
 write.csv(df, "data/dpc-augmented.csv")
 
 # test: check that no provinces weren't mapped to demographic data

### 2020-04-16 R/ClimateForcingModel.R
fix some more typos

@@ -25,7 +25,7 @@ df <- (
 )
 
 X <- abind(split(
-  select(df, c('date', 'air_temp', 'RH', 'gdppercapita', 'density')),
+  select(df, c('air_temp', 'RH', 'gdppercapita', 'density')),
   df$denominazione_provincia
 ), along=3)
 
@@ -47,7 +47,7 @@ cf_data <- list(
   T = nrow(I),
   w = w,
   i = I,
-  x = X
+  x = as.array(X)
 )
 
 # fit stan model

### 2020-04-16 R/ClimateForcingModel.R
add importations to model

@@ -53,3 +53,4 @@ cf_data <- list(
 # fit stan model
 cf_fit <- sampling(cf_mod, data = cf_data, iter = 400,
                    verbose = TRUE, chains = 4)
+

### 2020-04-16 R/ClimateForcingModel.R
Impose reasonable prior on beta

@@ -24,8 +24,19 @@ df <- (
   %>% ungroup
 )
 
+standardize <- function(series) {
+  (series - mean(series)) / sd(series)
+}
+df <- (
+  df
+  %>% mutate(std_air_temp = standardize(air_temp),
+             std_RH = standardize(RH),
+             std_gdppercapita = standardize(gdppercapita),
+             std_density = standardize(density))
+)
+
 X <- abind(split(
-  select(df, c('air_temp', 'RH', 'gdppercapita', 'density')),
+  select(df, c('std_air_temp', 'std_RH', 'std_gdppercapita', 'std_density')),
   df$denominazione_provincia
 ), along=3)
 
@@ -54,3 +65,4 @@ cf_data <- list(
 cf_fit <- sampling(cf_mod, data = cf_data, iter = 400,
                    verbose = TRUE, chains = 4)
 
+save(cf_fit, file="data/cf_fit.rds")

### 2020-04-16 R/ThompsonModel.R
moving files to R/

@@ -1,0 +1,52 @@
+#' Follows the Framework of the Thompson et al paper.
+#'
+
+library(EpiEstim)
+library(ggplot2)
+
+#' Let's use the Italian Provincial datasets.
+#'
+dpc_df <- read.csv("data/dpc-covid19-ita-province.csv")
+region_names <- unique(dpc_df$denominazione_provincia)
+all_res_parametric_si <- data.frame()
+
+
+augmented_df <- read.csv("data/dpc-augmented.csv")
+variables_df <- subset(augmented_df, select=c("denominazione_provincia", "gdppercapita", "density", "date", "air_temp", "dew_point"))
+
+
+
+for (region in region_names) {
+  region_df <- subset(dpc_df, denominazione_provincia == region, select = c("data", "totale_casi"))
+  names(region_df)[1] <- "date"
+  names(region_df)[2] <- "cases"
+  
+  #' Add a "new cases" column for per-day data.
+  region_df <- within(region_df, new_cases <- c(0, diff(cases)))
+  region_df$new_cases[region_df$new_cases<0] <- 0
+  region_df
+  
+  #' Analysis Step 1: Estimate the serial interval
+  #'
+  #' Let's just put these placeholders in here.
+  
+  est_mean_si = 2.5
+  est_std_si = 1
+  
+  #' Analysis Step 2: Estimate the reproduction rate.
+  #'
+  res_parametric_si <- estimate_R(region_df$new_cases,
+        method="parametric_si",
+        config = make_config(list(
+          mean_si = est_mean_si, 
+          std_si = est_std_si)))cs
+  
+  print(region_df$date)
+  
+  #' Analysis Step 3: Linear Regression Against Climate (and other confounding variables).
+  #'
+  res_parametric_si
+  break;
+}
+
+

### 2020-04-16 ThompsonModel.R

@@ -1,41 +1,0 @@
-#' Follows the Framework of the Thompson et al paper.
-#'
-
-library(EpiEstim)
-library(ggplot2)
-
-#' Let's use the Italian Provincial datasets.
-#'
-dpc_df <- read.csv("data/dpc-covid19-ita-province.csv")
-region_names <- unique(dpc_df$denominazione_provincia)
-
-
-#' Let's just try Rome first.
-#'
-region <- "Roma"
-region_df <- subset(dpc_df, denominazione_provincia == region, select = c("data", "totale_casi"))
-names(region_df)[1] <- "date"
-names(region_df)[2] <- "cases"
-
-#' Add a "new cases" column for per-day data. This might already be true.
-region_df <- within(region_df, new_cases <- c(0, diff(cases)))
-region_df
-
-#' Analysis Step 1: Estimate the serial interval
-#'
-#' Let's just put these placeholders in here.
-
-est_mean_si = 2.5
-est_std_si = 1
-
-#' Analysis Step 2: Estimate the reproduction rate.
-#'
-#'
-res_parametric_si <- estimate_R(region_df$new_cases,
-      method="parametric_si",
-      config = make_config(list(
-        mean_si = est_mean_si, 
-        std_si = est_std_si)))
-
-plot(res_parametric_si, legend = FALSE)
-

### 2020-04-16 R/ClimateForcingModel.R
Try with per capita importations and more iterations

@@ -24,6 +24,10 @@ df <- (
   %>% ungroup
 )
 
+# TODO: also kind of a hack, need to figure out the best actual date to stop the analysis
+# Using peak of reported cases as a hack
+df <- filter(df, date <= "2020-03-20")
+
 standardize <- function(series) {
   (series - mean(series)) / sd(series)
 }
@@ -42,6 +46,8 @@ X <- abind(split(
 
 I <- abind(split(df$infections, df$denominazione_provincia), along=2)
 
+pop <- abind(df %>% group_by(denominazione_provincia) %>% group_map(~ .x$population[1]), along=1)
+
 # TODO: use a correct discretized serial interval distribution
 w <- rep(0, nrow(I))
 w[1] <- 0.1
@@ -49,7 +55,9 @@ w[2] <- 0.2
 w[3] <- 0.3
 w[4] <- 0.2
 w[5] <- 0.1
-w[6] <- 0.1
+w[6] <- 0.05
+w[7] <- 0.03
+w[8] <- 0.02
 
 # stan data
 cf_data <- list(
@@ -58,11 +66,12 @@ cf_data <- list(
   T = nrow(I),
   w = w,
   i = I,
-  x = as.array(X)
+  x = as.array(X),
+  pop = pop
 )
 
 # fit stan model
-cf_fit <- sampling(cf_mod, data = cf_data, iter = 400,
+cf_fit <- sampling(cf_mod, data = cf_data, iter = 800,
                    verbose = TRUE, chains = 4)
 
 save(cf_fit, file="data/cf_fit.rds")

### 2020-04-16 R/ClimateForcingModel.R
minor tweaks

@@ -26,7 +26,7 @@ df <- (
 
 # TODO: also kind of a hack, need to figure out the best actual date to stop the analysis
 # Using peak of reported cases as a hack
-df <- filter(df, date <= "2020-03-20")
+df <- filter(df, date <= "2020-03-13")
 
 standardize <- function(series) {
   (series - mean(series)) / sd(series)
@@ -71,7 +71,7 @@ cf_data <- list(
 )
 
 # fit stan model
-cf_fit <- sampling(cf_mod, data = cf_data, iter = 800,
-                   verbose = TRUE, chains = 4)
+cf_fit <- sampling(cf_mod, data = cf_data, iter = 1000,
+                   verbose = TRUE, chains = 5)
 
 save(cf_fit, file="data/cf_fit.rds")

### 2020-04-16 R/ThompsonModel.R
italian column renaming

@@ -7,22 +7,18 @@ library(ggplot2)
 #' Let's use the Italian Provincial datasets.
 #'
 dpc_df <- read.csv("data/dpc-covid19-ita-province.csv")
-region_names <- unique(dpc_df$denominazione_provincia)
+province_names <- unique(dpc_df$province)
 all_res_parametric_si <- data.frame()
 
 
 augmented_df <- read.csv("data/dpc-augmented.csv")
-variables_df <- subset(augmented_df, select=c("denominazione_provincia", "gdppercapita", "density", "date", "air_temp", "dew_point"))
+variables_df <- subset(augmented_df, select=c("province", "gdppercapita", "density", "date", "air_temp", "dew_point"))
 
-
-
-for (region in region_names) {
-  region_df <- subset(dpc_df, denominazione_provincia == region, select = c("data", "totale_casi"))
-  names(region_df)[1] <- "date"
-  names(region_df)[2] <- "cases"
+for (province_name in province_names) {
+  region_df <- subset(dpc_df, province == province_name, select = c("date", "total_cases"))
   
   #' Add a "new cases" column for per-day data.
-  region_df <- within(region_df, new_cases <- c(0, diff(cases)))
+  region_df <- within(region_df, new_cases <- c(0, diff(total_cases)))
   region_df$new_cases[region_df$new_cases<0] <- 0
   region_df
   

### 2020-04-16 R/ThompsonModel.R
thompson model lm

@@ -8,11 +8,14 @@ library(ggplot2)
 #'
 dpc_df <- read.csv("data/dpc-covid19-ita-province.csv")
 province_names <- unique(dpc_df$province)
-all_res_parametric_si <- data.frame()
 
+all_provinces_df <- data.frame(matrix(ncol = 7, nrow = 0))
+colnames(all_provinces_df) <- c("date", "province", "mean_R", "gdppercapita","density","air_temp","dew_point")
 
 augmented_df <- read.csv("data/dpc-augmented.csv")
-variables_df <- subset(augmented_df, select=c("province", "gdppercapita", "density", "date", "air_temp", "dew_point"))
+variables_df <- subset(augmented_df, select=c("denominazione_provincia", "gdppercapita", "density", "date", "air_temp", "dew_point"))
+names(variables_df)[names(variables_df) == "denominazione_provincia"] <- "province"
+variables_df$date <- as.Date(variables_df$date, format = "%Y-%m-%d")
 
 for (province_name in province_names) {
   region_df <- subset(dpc_df, province == province_name, select = c("date", "total_cases"))
@@ -20,7 +23,7 @@ for (province_name in province_names) {
   #' Add a "new cases" column for per-day data.
   region_df <- within(region_df, new_cases <- c(0, diff(total_cases)))
   region_df$new_cases[region_df$new_cases<0] <- 0
-  region_df
+  region_df$date <- as.Date(region_df$date, format = "%Y-%m-%d")
   
   #' Analysis Step 1: Estimate the serial interval
   #'
@@ -35,14 +38,19 @@ for (province_name in province_names) {
         method="parametric_si",
         config = make_config(list(
           mean_si = est_mean_si, 
-          std_si = est_std_si)))cs
+          std_si = est_std_si)))
   
-  print(region_df$date)
+  #' res_parametric_si dataframe R that is based on aggregating on a weekly sliding window.
+  #' Therefore, the length of the date list and the mean list is off by 7.
+  l = length(region_df$date)
+  daily_R <- data.frame(date=region_df$date[1:(l-7)], mean_R=res_parametric_si$R$`Mean(R)`, province=province_name)
   
   #' Analysis Step 3: Linear Regression Against Climate (and other confounding variables).
   #'
-  res_parametric_si
-  break;
+  merged_df <- merge(daily_R, variables_df, by=c("date", "province"))
[…]

### 2020-04-16 R/ClimateForcingModel.R
Use 5 different stations

@@ -75,3 +75,6 @@ cf_fit <- sampling(cf_mod, data = cf_data, iter = 1000,
                    verbose = TRUE, chains = 5)
 
 save(cf_fit, file="data/cf_fit.rds")
+
+descr = c("temp", "RH", "GDP", "density")
+params <- rstan::extract(cf_fit)

### 2020-04-16 R/ForcingDataAugmentation.R

@@ -2,6 +2,7 @@ library(restatapi)
 library(dplyr)
 library(worldmet)
 library(memoise)
+library(safejoin)
 
 # Obtain mapping of health ministry province names to eurostat codes
 
@@ -91,28 +92,41 @@ augmentDPCdemo <- function(dpc, demodata) {
   df
 }
 
-closestweatherstation <- function(lat, long) {
-  getMeta(lat = lat, lon = long, n = 1)$code[1]
+closestweatherstations <- function(lat, long, n) {
+  getMeta(lat = lat, lon = long, n = n)$code
 }
-closestweatherstationM <- memoise(closestweatherstation)
+closestweatherstationsM <- memoise(closestweatherstations)
 
 importNOAAM <- memoise(importNOAA)
 
-# Augment a single group of data points with weather, using just one station code.
+# Augment a single group of data points with weather, using a list of station codes.
 augmentsingleweather <- function(group, station) {
-  print(station$station[1])
-  noaa <- importNOAAM(code = station$station[1], year = 2020)
-  print(group)
-  print(noaa)
-  result <- left_join(group, noaa, by = "date")
-  print(result)
+  fallbacks <- strsplit(station$station[1], ",")
+  result <- group
+  for (code in fallbacks) {
+    print(paste("Looking up weather for ", code))
+    noaa <- importNOAAM(code = code, year = 2020)
+    result <- safe_left_join(result, noaa, by = "date", conflict = coalesce)
+    if (all(!is.na(result$air_temp)) & all(!is.na(result$RH))) {
+      break;
+    }
+  }
   select(result, -c("station"))
 }
 
 # Augment a table with date, lat, and long columns with weather data collected closest to the given time.
 augmentDPCweather <- function(dpc) {
   # 1. Find the appropriate station code for each latitude and longitude present in data.
-  dpc <- mutate(dpc, station = mapply(closestweatherstationM, lat = lat, long = long))
+  stations = mapply(
+    closestweatherstationsM,
+    lat = dpc$lat,
+    long = dpc$long,
+    MoreArgs = list(n = 5)
+  )
+  dpc <- mutate(
+    dpc,
+    station = apply(stations, 2, function (sts) { paste(sts, collapse=",", sep=",") })
+  )
   dpc <- mutate(dpc, date = as.POSIXct(data, format="%Y-%m-%dT%H:%M:%S", tz = "Europe/Rome"))
   dpc <- select(dpc, -c("data"))
[…]

### 2020-04-16 R/ClimateForcingModel.R
Drop last date a little earlier

@@ -13,10 +13,7 @@ df <- mutate(
   df,
   infections = coalesce(pmax(0, totale_casi - lag(totale_casi)), 0)
 )
-# TODO: this is a hack, do this a better way...
-# Drop the last date as it is NA for all locations.
-# Then drop the locations with some NA in data.
-df <- filter(df, date != max(as.character(date)))
+# Drop the locations with some NA in data (there should be none, TODO delete this code).
 df <- (
   df
   %>% group_by(denominazione_provincia)

### 2020-04-16 R/ForcingDataAugmentation.R

@@ -2,7 +2,6 @@ library(restatapi)
 library(dplyr)
 library(worldmet)
 library(memoise)
-library(safejoin)
 
 # Obtain mapping of health ministry province names to eurostat codes
 
@@ -101,17 +100,13 @@ importNOAAM <- memoise(importNOAA)
 
 # Augment a single group of data points with weather, using a list of station codes.
 augmentsingleweather <- function(group, station) {
-  fallbacks <- strsplit(station$station[1], ",")
-  result <- group
-  for (code in fallbacks) {
-    print(paste("Looking up weather for ", code))
-    noaa <- importNOAAM(code = code, year = 2020)
-    result <- safe_left_join(result, noaa, by = "date", conflict = coalesce)
-    if (all(!is.na(result$air_temp)) & all(!is.na(result$RH))) {
-      break;
-    }
-  }
-  select(result, -c("station"))
+  fallbacks <- strsplit(station$station[1], ",")[[1]]
+  print(paste("Looking up weather for", fallbacks))
+  noaa <- importNOAAM(code = fallbacks, year = 2020)
+  noaa <- noaa %>% group_by(date) %>% summarize(
+    air_temp = first(na.omit(air_temp)), RH = first(na.omit(RH))
+  )
+  left_join(group, noaa, by = "date")
 }
 
 # Augment a table with date, lat, and long columns with weather data collected closest to the given time.
@@ -133,11 +128,14 @@ augmentDPCweather <- function(dpc) {
 }
 
 # Repair total case data to be monotonically increasing, through taking a rolling maximum.
+# Also drop the last date as it is likely to have no weather data for each location.
 repairtotalcases <- function(dpc) {
   (dpc
    %>% group_by(denominazione_provincia)
    %>% group_modify(~ arrange(.x, by = data) %>% mutate(totale_casi = cummax(totale_casi)))
-   %>% ungroup)
+   %>% ungroup
+   %>% filter(data != max(as.character(data)))
+   )
 }
 
 # test code
@@ -152,5 +150,5 @@ write.csv(df, "data/dpc-augmented.csv")
 # test: check that no provinces weren't mapped to demographic data
 stopifnot(nrow(filter(df, is.na(density))) == 0)
 
-# test: check no temperature NAs (fails! TODO)
+# test: check no temperature NAs
 stopifnot(nrow(filter(df, is.na(air_temp))) == 0)

### 2020-04-16 R/ClimateForcingModel.R
update model

@@ -68,8 +68,7 @@ cf_data <- list(
 )
 
 # fit stan model
-cf_fit <- sampling(cf_mod, data = cf_data, iter = 1000,
-                   verbose = TRUE, chains = 5)
+cf_fit <- sampling(cf_mod, data = cf_data, verbose = TRUE)
 
 save(cf_fit, file="data/cf_fit.rds")
 

### 2020-04-16 R/ClimateForcingModel.R
Use correct SI distribution

@@ -1,19 +1,24 @@
 library(rstan)
 library(dplyr)
+library(tidyr)
 library(abind)
 
-# compile stan model
+# Compile stan model
 cf_mod <- stan_model("ClimateForcingModel.stan")
-
 save(cf_mod, file="data/cf_mod.rds")
 
 df <- read.csv("data/dpc-augmented.csv", stringsAsFactors = FALSE)
-# TODO: this is probably duplicated work from somewhere
+
+# TODO:
+# This is probably duplicated work from the Thompson model. Try to merge the
+# codepaths that generate this df.
 df <- mutate(
   df,
   infections = coalesce(pmax(0, totale_casi - lag(totale_casi)), 0)
 )
-# Drop the locations with some NA in data (there should be none, TODO delete this code).
+
+# Drop the locations with some NA in data (there should be none, TODO: delete
+# this code).
 df <- (
   df
   %>% group_by(denominazione_provincia)
@@ -21,10 +26,12 @@ df <- (
   %>% ungroup
 )
 
-# TODO: also kind of a hack, need to figure out the best actual date to stop the analysis
-# Using peak of reported cases as a hack
+# Stop the analysis on Mar 13
 df <- filter(df, date <= "2020-03-13")
 
+# Standardize regressors.
+# TODO: Remember the original scale of the series so that the model coefficients
+# are interpretable.
 standardize <- function(series) {
   (series - mean(series)) / sd(series)
 }
@@ -36,6 +43,7 @@ df <- (
              std_density = standardize(density))
 )
 
+# Generate data for climate model.
 X <- abind(split(
   select(df, c('std_air_temp', 'std_RH', 'std_gdppercapita', 'std_density')),
   df$denominazione_provincia
@@ -45,16 +53,20 @@ I <- abind(split(df$infections, df$denominazione_provincia), along=2)
 
 pop <- abind(df %>% group_by(denominazione_provincia) %>% group_map(~ .x$population[1]), along=1)
 
-# TODO: use a correct discretized serial interval distribution
-w <- rep(0, nrow(I))
-w[1] <- 0.1
-w[2] <- 0.2
-w[3] <- 0.3
-w[4] <- 0.2
-w[5] <- 0.1
-w[6] <- 0.05
-w[7] <- 0.03
-w[8] <- 0.02
+#' Discretize the serial interval distribution of 4.7 std. dev 2.9, used in the
+#' CMMID paper https://epiforecasts.io/covid/methods.html
[…]

### 2020-04-16 R/Plots.R

@@ -1,0 +1,38 @@
+library(bayesplot)
+library(ggplot2)
+
+load("data/cf_fit.rds")
+
+descr = c("temp", "RH", "GDP per capita", "density")
+params <- rstan::extract(cf_fit)
+
+mcmc_areas(
+  cf_fit,
+  regex_pars = c("beta.+"),
+  point_est = "mean"
+)
+
+mcmc_areas(
+  cf_fit,
+  regex_pars = c("theta\\[.+"),
+  point_est = "mean",
+  transformations = exp
+)
+
+# graph 1: histogram of beta
+par(mfrow=c(3, 2))
+for (i in 1:4) {
+  hist(params$beta[,i], main = descr[i], xlab = paste("coef for GDP per capita", descr[i]))
+}
+
+# graph 2: evolution of climate-independent R over time, nationally
+# Code copied from https://stackoverflow.com/a/51993331/3575047 written by @www.
+mean_cl_quantile <- function(x, q = c(0.1, 0.9), na.rm = TRUE) {
+  dat <- data.frame(y = mean(x, na.rm = na.rm),
+                    ymin = quantile(x, probs = q[1], na.rm = na.rm),
+                    ymax = quantile(x, probs = q[2], na.rm = na.rm))
+  return(dat)
+}
+thetadf <- data.frame(params$theta)
+thetadf <- pivot_longer(thetadf, cols = colnames(thetadf), names_to = "time", values_to = "logR")
+ggplot(thetadf, aes(time, logR)) + geom_smooth(stat = 'summary', fun.data = mean_cl_quantile)

### 2020-04-16 R/ClimateForcingModel.R
estimated R in certain location by date

@@ -83,3 +83,4 @@ cf_data <- list(
 cf_fit <- sampling(cf_mod, data = cf_data, verbose = TRUE)
 
 save(cf_fit, file="data/cf_fit.rds")
+save(cf_data, file="data/cf_data.rds")

### 2020-04-16 R/Plots.R

@@ -26,6 +26,7 @@ for (i in 1:4) {
 }
 
 # graph 2: evolution of climate-independent R over time, nationally
+# TODO: fix axis
 # Code copied from https://stackoverflow.com/a/51993331/3575047 written by @www.
 mean_cl_quantile <- function(x, q = c(0.1, 0.9), na.rm = TRUE) {
   dat <- data.frame(y = mean(x, na.rm = na.rm),
@@ -34,5 +35,16 @@ mean_cl_quantile <- function(x, q = c(0.1, 0.9), na.rm = TRUE) {
   return(dat)
 }
 thetadf <- data.frame(params$theta)
-thetadf <- pivot_longer(thetadf, cols = colnames(thetadf), names_to = "time", values_to = "logR")
-ggplot(thetadf, aes(time, logR)) + geom_smooth(stat = 'summary', fun.data = mean_cl_quantile)
+thetadf <- pivot_longer(thetadf, cols = colnames(thetadf), names_to = "date", values_to = "logR")
+(ggplot(thetadf, aes(date, exp(logR), group=1))
+ + geom_smooth(stat = 'summary', fun.data = mean_cl_quantile)
+ + ylab("R")
+ + ggtitle("Climate-independent Italian basic R, by date"))
+
+# graph 3: R in province of Roma over time
+thetadf <- data.frame(params$theta) + params$lambda[85] + params$epsilon[,85,] + params$beta %*% t(cf_data$x[,,85])
+thetadf <- pivot_longer(thetadf, cols = colnames(thetadf), names_to = "date", values_to = "logR")
+(ggplot(thetadf, aes(date, exp(logR), group=1))
+  + geom_smooth(stat = 'summary', fun.data = mean_cl_quantile)
+  + ylab("R")
+  + ggtitle("Estimated R in Rome, by date"))

### 2020-04-17 R/ClimateForcingModel.R
Update plots

@@ -27,14 +27,15 @@ df <- (
 )
 
 # Stop the analysis on Mar 13
-df <- filter(df, date <= "2020-03-13")
+df <- filter(df, date < "2020-03-14")
 
-# Standardize regressors.
-# TODO: Remember the original scale of the series so that the model coefficients
-# are interpretable.
+#' Standardize a single feature.
+#' @param series The series to standardize.
+#' @return A series with mean 0 and standard deviation 1.
 standardize <- function(series) {
   (series - mean(series)) / sd(series)
 }
+
 df <- (
   df
   %>% mutate(std_air_temp = standardize(air_temp),

### 2020-04-17 R/DataPipeline.R
worked on data pipeline

@@ -1,6 +1,17 @@
+library(restatapi)
+library(dplyr)
+library(worldmet)
+library(memoise)
 library(lubridate)
 
-#' Fetches a raw file from url.
+#' Fetches a raw file from url, and saves it to a filename.
+#' For proper style, filename string should begin with "data/".
+#' Throws exception on failure.
+#' 
+#' @param url The URL of the file to download.
+#' @param filename The filename that the file is saved to.
+#'
+#' @return None
 #'
 fetch_file <- function(url, filename) {
   tryCatch({
@@ -15,8 +26,10 @@ fetch_file <- function(url, filename) {
   })
 }
 
-#' Fetches the latest COVID-19 cases data from ECDC.
-#'
+#' Fetches the latest COVID-19 cases data from European Center for Disease Control (ECDC).
+#' 
+#' @returns None
+#' 
 fetch_latest_ecdc <- function() {
   url <- "https://opendata.ecdc.europa.eu/covid19/casedistribution/csv"
   filename <- "data/ecdc-COVID-19-up-to-date.csv"
@@ -33,10 +46,11 @@ fetch_latest_ecdc <- function() {
   names(d)[names(d) == "dateRep"] <- "DateRep"
   write.csv(d, filename)
 }
-fetch_latest_ecdc()
 
 #' Fetches latest data from John Hopkins University CSSE.
 #'
+#' @return None
+#' 
 fetch_latest_csse <- function() {
   url <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv"
   filename <- "data/time_series_covid19_confirmed_global.csv"
@@ -50,9 +64,11 @@ fetch_latest_csse <- function() {
   filename <- "data/time_series_covid19_recovered_global.csv"
   fetch_file(url, filename)
 }
-fetch_latest_csse()
 
-#' Fetch Italian provinces data.
+
+#' Fetch Italian provinces data from DPC.
+#' 
+#' @return None
 #'
 fetch_latest_dpc <- function() {
   url <- "https://raw.githubusercontent.com/pcm-dpc/COVID-19/master/dati-province/dpc-covid19-ita-province.csv"
@@ -69,4 +85,163 @@ fetch_latest_dpc <- function() {
   names(d)[names(d) == "totale_casi"] <- "total_cases"
[…]

### 2020-04-17 R/ForcingDataAugmentation.R

@@ -1,154 +1,0 @@
-library(restatapi)
-library(dplyr)
-library(worldmet)
-library(memoise)
-
-# Obtain mapping of health ministry province names to eurostat codes
-
-getprovincelist <- function() {
-  # The island of Sardinia was reorganized between collection of the demographic data and
-  # the health ministry's collection of COVID-19 data, so demographic data does not map
-  # cleanly. Because Sardinia is an island and the situation may be unique, we have just
-  # excluded it from the analysis.
-  sardiniaprefix <- "ITG2"
-
-  dsd <- get_eurostat_dsd("nama_10r_3gdp")
-  itprovinces <- select(
-    filter(
-      filter(dsd, concept == "GEO"),
-      startsWith(code, "IT") &
-        nchar(code) == 5 &
-        code != "ITZZZ" &
-        !startsWith(code, sardiniaprefix)
-    ),
-    c("code", "name")
-  )
-  # 3 provinces have names which do not match the ones the health ministry uses;
-  # so we rename them to match
-  statname <- c("Valle d'Aosta/Vallée d'Aoste", "Bolzano-Bozen", "Massa-Carrara")
-  healthname <- c("Aosta", "Bolzano", "Massa Carrara")
-  itprovinces <- left_join(
-    itprovinces,
-    data.frame(statname, healthname, stringsAsFactors = FALSE),
-    by = c("name" = "statname")
-  )
-  itprovinces$name <- coalesce(itprovinces$healthname, itprovinces$name)
-  select(itprovinces, c("code", "name"))
-}
-
-# Fetch GDP per capita data and population density data for Italian provinces
-# Source: Eurostat
-getdemodata <- function() {
-  itprovinces <- getprovincelist()
-
-  gdppercapita <- filter(get_eurostat_data(
-    "nama_10r_3gdp",
-    date_filter="2017",
-    filters=itprovinces$code
-  ), unit == "EUR_HAB")
-  gdppercapita <- select(gdppercapita, c("geo", "values"))
-  gdppercapita$geo <- as.character(gdppercapita$geo)
-
-  density <- get_eurostat_data(
-    "demo_r_d3dens",
-    date_filter="2017",
-    filters=itprovinces$code
-  )
-  density <- select(density, c("geo", "values"))
-  density$geo <- as.character(density$geo)
-
-  population <- filter(get_eurostat_data(
-    "demo_r_pjangrp3",
-    date_filter="2017",
[…]

### 2020-04-17 R/ThompsonModel.R

@@ -10,10 +10,10 @@ dpc_df <- read.csv("data/dpc-covid19-ita-province.csv")
 province_names <- unique(dpc_df$province)
 
 all_provinces_df <- data.frame(matrix(ncol = 7, nrow = 0))
-colnames(all_provinces_df) <- c("date", "province", "mean_R", "gdppercapita","density","air_temp","dew_point")
+colnames(all_provinces_df) <- c("date", "province", "mean_R", "gdppercapita", "density", "air_temp","RH")
 
 augmented_df <- read.csv("data/dpc-augmented.csv")
-variables_df <- subset(augmented_df, select=c("denominazione_provincia", "gdppercapita", "density", "date", "air_temp", "dew_point"))
+variables_df <- subset(augmented_df, select=c("denominazione_provincia", "gdppercapita", "density", "date", "air_temp", "RH"))
 names(variables_df)[names(variables_df) == "denominazione_provincia"] <- "province"
 variables_df$date <- as.Date(variables_df$date, format = "%Y-%m-%d")
 
@@ -27,10 +27,10 @@ for (province_name in province_names) {
   
   #' Analysis Step 1: Estimate the serial interval
   #'
-  #' Let's just put these placeholders in here.
-  
-  est_mean_si = 2.5
-  est_std_si = 1
+  #' We have determined that it is difficult to find the line-list data for COVID-19, and thus will be using
+  #' the mean and standard deviation 
+  est_mean_si = 4.7
+  est_std_si = 2.9
   
   #' Analysis Step 2: Estimate the reproduction rate.
   #'
@@ -51,6 +51,10 @@ for (province_name in province_names) {
   all_provinces_df <- rbind(all_provinces_df, merged_df)
 }
 
-all_provinces_df
-fit <- lm(mean_R ~ gdppercapita + density + air_temp + dew_point, data=all_provinces_df)
-summary(fit) # show results
+#' Cutoff on 2020-03-13, since intervention completely changes regime.
+all_provinces_df <- subset(all_provinces_df, date <= "2020-03-13")
+fit <- lm(mean_R ~ gdppercapita + density + air_temp + RH, data=all_provinces_df)
+summary(fit)
+
+# Produce good plots for the 
+#

### 2020-04-17 R/DataPipeline.R
Update plots

@@ -7,7 +7,7 @@ library(lubridate)
 #' Fetches a raw file from url, and saves it to a filename.
 #' For proper style, filename string should begin with "data/".
 #' Throws exception on failure.
-#' 
+#'
 #' @param url The URL of the file to download.
 #' @param filename The filename that the file is saved to.
 #'
@@ -27,14 +27,14 @@ fetch_file <- function(url, filename) {
 }
 
 #' Fetches the latest COVID-19 cases data from European Center for Disease Control (ECDC).
-#' 
+#'
 #' @returns None
-#' 
+#'
 fetch_latest_ecdc <- function() {
   url <- "https://opendata.ecdc.europa.eu/covid19/casedistribution/csv"
   filename <- "data/ecdc-COVID-19-up-to-date.csv"
   fetch_file(url, filename)
-  
+
   #' Quick cleanup on the column titles.
   #'
   d <- read.csv(filename, stringsAsFactors = FALSE)
@@ -50,16 +50,16 @@ fetch_latest_ecdc <- function() {
 #' Fetches latest data from John Hopkins University CSSE.
 #'
 #' @return None
-#' 
+#'
 fetch_latest_csse <- function() {
   url <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv"
   filename <- "data/time_series_covid19_confirmed_global.csv"
   fetch_file(url, filename)
-  
+
   url <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv"
   filename <- "data/time_series_covid19_deaths_global.csv"
   fetch_file(url, filename)
-  
+
   url <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_recovered_global.csv"
   filename <- "data/time_series_covid19_recovered_global.csv"
   fetch_file(url, filename)
@@ -67,14 +67,14 @@ fetch_latest_csse <- function() {
 
 
 #' Fetch Italian provinces data from DPC.
-#' 
+#'
 #' @return None
 #'
 fetch_latest_dpc <- function() {
[…]

### 2020-04-17 R/Plots.R

@@ -6,7 +6,10 @@ load("data/cf_fit.rds")
 
 descr = c("temp", "RH", "GDP per capita", "density")
 params <- rstan::extract(cf_fit)
-dates <- c(paste("02-", 24:29, sep=""), paste("03-0", 1:9, sep=""), paste("03-", 10:13, sep=""))
+dates <-
+  c(paste("02-", 24:29, sep = ""),
+    paste("03-0", 1:9, sep = ""),
+    paste("03-", 10:13, sep = ""))
 
 #' Given the Bayesian posterior stanfit object, plot the MCMC confidence for
 #' beta.
@@ -15,11 +18,14 @@ dates <- c(paste("02-", 24:29, sep=""), paste("03-0", 1:9, sep=""), paste("03-",
 #' @return A plot of location of standardized beta parameters, with 50%
 #'   confidence interval highlighted.
 plot_beta_confidence <- function(cf_fit) {
-  mcmc_areas(
-    cf_fit,
-    regex_pars = c("beta.+"),
-    point_est = "mean"
-  ) + ggtitle("Posterior location of β (magnitude of climate forcings)")
+  (
+    mcmc_areas(
+      cf_fit,
+      regex_pars = c("beta.+"),
+      point_est = "mean"
+    ) + ggplot2::labs(title = "Posterior of β, magnitude of climate forcings",
+                      subtitle = "Each standardized; in order: temp., rel. humidity, GDP p.c., density")
+  )
 }
 
 #' Given the Bayesian posterior stanfit object, plot the MCMC confidence for
@@ -29,17 +35,22 @@ plot_beta_confidence <- function(cf_fit) {
 #' @return A plot of location of standardized beta parameters, with 50%
 #'   confidence interval highlighted.
 plot_lambda_confidence <- function(cf_fit, start, end) {
-  mcmc_intervals(
-    cf_fit,
-    pars = paste("lambda[", start:end, "]", sep=""),
-    point_est = "mean"
-  ) + ggplot2::labs(
-    title = "Posterior of λ, location-dependent adjustment to log(R)",
-    subtitle = paste("Mean, 50% and 90% confidence intervals for locations", start, "through", end),
-    x = "Local adjustment to log(R)"
-  ) + ggplot2::theme(
-    axis.text.y = element_blank(),
-    axis.ticks.y = element_blank()
+  (
+    mcmc_intervals(
+      cf_fit,
+      pars = paste("lambda[", start:end, "]", sep = ""),
[…]

### 2020-04-17 R/Plots.R
Add plots

@@ -92,7 +92,10 @@ plot_independentR <- function(params, cf_data) {
     ggplot(thetadf, aes(date, exp(logR), group = 1))
     + geom_smooth(stat = 'summary', fun.data = mean_cl_quantile)
     + ylab("R")
-    + ggtitle("Climate-independent basic R, by date")
+    + ggplot2::labs(
+      title = "Climate-independent basic R, by date",
+      subtitle = "Mean and 80% confidence interval"
+    )
     + theme(axis.text.x = element_text(angle = 90))
   )
 }
@@ -122,7 +125,10 @@ plot_locationR <- function(params, cf_data, loc, locname) {
     ggplot(thetadf, aes(date, exp(logR), group = 1))
     + geom_smooth(stat = 'summary', fun.data = mean_cl_quantile)
     + ylab("R")
-    + ggtitle(paste("Estimated R in", locname, "by date"))
+    + ggplot2::labs(
+      title = paste("Estimated R in", locname, "by date"),
+      subtitle = "Mean and 80% confidence interval"
+    )
     + theme(axis.text.x = element_text(angle = 90))
   )
 }
@@ -147,4 +153,5 @@ plot_lambda_confidence(cf_fit, 69, 102)
 plot_independentR(params, cf_data)
 plot_locationR(params, cf_data, 13, "Bergamo")
 plot_locationR(params, cf_data, 79, "Rome")
+plot_locationR(params, cf_data, 96, "Venezia")
 plot_epsilon_dispersion(params)

### 2020-04-17 R/Plots.R
update plots

@@ -40,7 +40,8 @@ plot_lambda_confidence <- function(cf_fit, start, end) {
       cf_fit,
       pars = paste("lambda[", start:end, "]", sep = ""),
       point_est = "mean"
-    ) + ggplot2::labs(
+    )
+    + ggplot2::labs(
       title = "Posterior of λ, location-dependent adjustment to log(R)",
       subtitle = paste(
         "Mean, 50% and 90% confidence intervals for locations",
@@ -49,8 +50,10 @@ plot_lambda_confidence <- function(cf_fit, start, end) {
         end
       ),
       x = "Local adjustment to log(R)"
-    ) + ggplot2::theme(axis.text.y = element_blank(),
-                       axis.ticks.y = element_blank())
+    )
+    + ggplot2::theme(axis.text.y = element_blank(),
+                     axis.ticks.y = element_blank())
+    + xlim(-2, 2)
   )
 }
 
@@ -152,6 +155,8 @@ plot_lambda_confidence(cf_fit, 35, 68)
 plot_lambda_confidence(cf_fit, 69, 102)
 plot_independentR(params, cf_data)
 plot_locationR(params, cf_data, 13, "Bergamo")
+plot_locationR(params, cf_data, 55, "Milano")
 plot_locationR(params, cf_data, 79, "Rome")
 plot_locationR(params, cf_data, 96, "Venezia")
+plot_locationR(params, cf_data, 99, "Verona")
 plot_epsilon_dispersion(params)

### 2020-04-17 R/DataPipeline.R
data pipeline migration done

@@ -92,6 +92,7 @@ fetch_latest_csse()
 
 #' Obtain mapping of health ministry province names to eurostat codes
 #'
+#' @return Dataframe of italian provinces with columns "name", "code".
 getprovincelist <- function() {
   # The island of Sardinia was reorganized between collection of the demographic data and
   # the health ministry's collection of COVID-19 data, so demographic data does not map
@@ -126,27 +127,32 @@ getprovincelist <- function() {
 #' Fetch GDP per capita data and population density data for Italian provinces
 #' Source: Eurostat
 #'
+#' @return Eurostat dataframe with columns "name", "gdppercapita", "density", "population".
 getdemodata <- function() {
   itprovinces <- getprovincelist()
   
+  #' Get GDP/capita data from Eurostat.
+  #'
   gdppercapita <- filter(get_eurostat_data(
-    "nama_10r_3gdp",
+    "nama_10r_3gdp", # Specific column code for GDP per capital from Eurostat documentation.
     date_filter="2017",
     filters=itprovinces$code
   ), unit == "EUR_HAB")
   gdppercapita <- select(gdppercapita, c("geo", "values"))
   gdppercapita$geo <- as.character(gdppercapita$geo)
   
+  #' Get population density data from Eurostat.
   density <- get_eurostat_data(
-    "demo_r_d3dens",
+    "demo_r_d3dens", # Specific column code for density from Eurostat documentation.
     date_filter="2017",
     filters=itprovinces$code
   )
   density <- select(density, c("geo", "values"))
   density$geo <- as.character(density$geo)
   
+  #' Get population data from Eurostat.
   population <- filter(get_eurostat_data(
-    "demo_r_pjangrp3",
+    "demo_r_pjangrp3", # Specific column code for population from Eurostat documentation.
     date_filter="2017",
     filters=itprovinces$code
   ), sex == "T" & age == "TOTAL")
@@ -167,31 +173,45 @@ getdemodata <- function() {
   full_join(df, itprovinces, by = c("geo" = "code"))
 }
 
-# Augment Italy DPC data with demographics
-# Also remove Sardinia provinces
+#' Augment Italy DPC data with demographics, and also removes Sardinia provinces
+#' 
+#' @param dpc Italian per province timeseries data.
+#' @param demodata Dataframe sourced from Eurostat that contains GDP per Capita and Density information on Italian regions.
+#' 
+#' @return Augmented dataframe that contains both COVID-19 cases timeseries and eurostats.
[…]

### 2020-04-17 R/ThompsonModel.R

@@ -13,8 +13,7 @@ all_provinces_df <- data.frame(matrix(ncol = 7, nrow = 0))
 colnames(all_provinces_df) <- c("date", "province", "mean_R", "gdppercapita", "density", "air_temp","RH")
 
 augmented_df <- read.csv("data/dpc-augmented.csv")
-variables_df <- subset(augmented_df, select=c("denominazione_provincia", "gdppercapita", "density", "date", "air_temp", "RH"))
-names(variables_df)[names(variables_df) == "denominazione_provincia"] <- "province"
+variables_df <- subset(augmented_df, select=c("province", "gdppercapita", "density", "date", "air_temp", "RH"))
 variables_df$date <- as.Date(variables_df$date, format = "%Y-%m-%d")
 
 for (province_name in province_names) {

### 2020-04-17 R/ClimateForcingModel.R
add diff to Data pipeline

@@ -9,19 +9,11 @@ save(cf_mod, file="data/cf_mod.rds")
 
 df <- read.csv("data/dpc-augmented.csv", stringsAsFactors = FALSE)
 
-# TODO:
-# This is probably duplicated work from the Thompson model. Try to merge the
-# codepaths that generate this df.
-df <- mutate(
-  df,
-  infections = coalesce(pmax(0, totale_casi - lag(totale_casi)), 0)
-)
-
 # Drop the locations with some NA in data (there should be none, TODO: delete
 # this code).
 df <- (
   df
-  %>% group_by(denominazione_provincia)
+  %>% group_by(province)
   %>% filter(!any(is.na(air_temp)) & !any(is.na(RH)))
   %>% ungroup
 )
@@ -50,9 +42,9 @@ X <- abind(split(
   df$denominazione_provincia
 ), along=3)
 
-I <- abind(split(df$infections, df$denominazione_provincia), along=2)
+I <- abind(split(df$new_cases, df$province), along=2)
 
-pop <- abind(df %>% group_by(denominazione_provincia) %>% group_map(~ .x$population[1]), along=1)
+pop <- abind(df %>% group_by(province) %>% group_map(~ .x$population[1]), along=1)
 
 #' Discretize the serial interval distribution of 4.7 std. dev 2.9, used in the
 #' CMMID paper https://epiforecasts.io/covid/methods.html

### 2020-04-17 R/DataPipeline.R

@@ -239,7 +239,7 @@ augmentDPCweather <- function(dpc) {
     dpc,
     station = apply(stations, 2, function (sts) { paste(sts, collapse=",", sep=",") })
   )
-  dpc <- mutate(dpc, date = as.POSIXct("date", format="%Y-%m-%dT%H:%M:%S", tz = "Europe/Rome"))
+  dpc <- mutate(dpc, date = as.POSIXct(date, format="%Y-%m-%dT%H:%M:%S", tz = "Europe/Rome"))
   group_by(dpc, station) %>% group_modify(augmentsingleweather) %>% ungroup
 }
 
@@ -253,10 +253,12 @@ repairtotalcases <- function(dpc) {
   (dpc
    %>% group_by(province)
    %>% group_modify(~ arrange(.x, by=date) %>% mutate(total_cases = cummax(total_cases)))
+   %>% mutate(new_cases = c(0, diff(total_cases)))
    %>% ungroup
    %>% filter(date != max(as.character(date)))
   )
 }
+
 
 #' Code which augments the Italian per province data with GDP, weather, density data.
 #' 

### 2020-04-17 R/ThompsonModel.R

@@ -6,23 +6,15 @@ library(ggplot2)
 
 #' Let's use the Italian Provincial datasets.
 #'
-dpc_df <- read.csv("data/dpc-covid19-ita-province.csv")
+dpc_df <- read.csv("data/dpc-augmented.csv")
+dpc_df$date <- as.Date(dpc_df$date, format = "%Y-%m-%d")
 province_names <- unique(dpc_df$province)
 
 all_provinces_df <- data.frame(matrix(ncol = 7, nrow = 0))
 colnames(all_provinces_df) <- c("date", "province", "mean_R", "gdppercapita", "density", "air_temp","RH")
 
-augmented_df <- read.csv("data/dpc-augmented.csv")
-variables_df <- subset(augmented_df, select=c("province", "gdppercapita", "density", "date", "air_temp", "RH"))
-variables_df$date <- as.Date(variables_df$date, format = "%Y-%m-%d")
-
 for (province_name in province_names) {
-  region_df <- subset(dpc_df, province == province_name, select = c("date", "total_cases"))
-  
-  #' Add a "new cases" column for per-day data.
-  region_df <- within(region_df, new_cases <- c(0, diff(total_cases)))
-  region_df$new_cases[region_df$new_cases<0] <- 0
-  region_df$date <- as.Date(region_df$date, format = "%Y-%m-%d")
+  region_df <- subset(dpc_df, province == province_name)
   
   #' Analysis Step 1: Estimate the serial interval
   #'
@@ -46,14 +38,11 @@ for (province_name in province_names) {
   
   #' Analysis Step 3: Linear Regression Against Climate (and other confounding variables).
   #'
-  merged_df <- merge(daily_R, variables_df, by=c("date", "province"))
-  all_provinces_df <- rbind(all_provinces_df, merged_df)
+  merged_df <- merge(daily_R, region_df, by=c("date", "province"))
+  all_provinces_df <- rbind(all_provinces_df, subset(merged_df, select=c("date", "province", "mean_R", "gdppercapita", "density", "air_temp","RH")))
 }
 
 #' Cutoff on 2020-03-13, since intervention completely changes regime.
 all_provinces_df <- subset(all_provinces_df, date <= "2020-03-13")
 fit <- lm(mean_R ~ gdppercapita + density + air_temp + RH, data=all_provinces_df)
 summary(fit)
-
-# Produce good plots for the 
-#

### 2020-04-17 R/ClimateForcingModel.R
fixed small bug in forcing climate model

@@ -9,15 +9,6 @@ save(cf_mod, file="data/cf_mod.rds")
 
 df <- read.csv("data/dpc-augmented.csv", stringsAsFactors = FALSE)
 
-# Drop the locations with some NA in data (there should be none, TODO: delete
-# this code).
-df <- (
-  df
-  %>% group_by(province)
-  %>% filter(!any(is.na(air_temp)) & !any(is.na(RH)))
-  %>% ungroup
-)
-
 # Stop the analysis on Mar 13
 df <- filter(df, date < "2020-03-14")
 
@@ -39,7 +30,7 @@ df <- (
 # Generate data for climate model.
 X <- abind(split(
   select(df, c('std_air_temp', 'std_RH', 'std_gdppercapita', 'std_density')),
-  df$denominazione_provincia
+  df$province
 ), along=3)
 
 I <- abind(split(df$new_cases, df$province), along=2)

### 2020-04-17 R/DataPipeline.R

@@ -223,6 +223,7 @@ augmentsingleweather <- function(group, station) {
 }
 
 #' Augment a table with date, lat, and long columns with weather data collected closest to the given time.
+#' Warning: This function will be very slow, since it queries the NOAA database often.
 #'
 #' @param dpc Dataframe of DPC per province time series case data. Required columns: "date", "lat", "long".
 #' 

### 2020-04-17 R/ThompsonModel.R

@@ -35,13 +35,12 @@ for (province_name in province_names) {
   #' Therefore, the length of the date list and the mean list is off by 7.
   l = length(region_df$date)
   daily_R <- data.frame(date=region_df$date[1:(l-7)], mean_R=res_parametric_si$R$`Mean(R)`, province=province_name)
-  
-  #' Analysis Step 3: Linear Regression Against Climate (and other confounding variables).
-  #'
   merged_df <- merge(daily_R, region_df, by=c("date", "province"))
   all_provinces_df <- rbind(all_provinces_df, subset(merged_df, select=c("date", "province", "mean_R", "gdppercapita", "density", "air_temp","RH")))
 }
 
+#' Analysis Step 3: Linear Regression Against Climate (and other confounding variables).
+#'
 #' Cutoff on 2020-03-13, since intervention completely changes regime.
 all_provinces_df <- subset(all_provinces_df, date <= "2020-03-13")
 fit <- lm(mean_R ~ gdppercapita + density + air_temp + RH, data=all_provinces_df)

### 2020-04-17 R/DataPipeline.R
Make a robust version

@@ -130,7 +130,7 @@ getprovincelist <- function() {
 #' @return Eurostat dataframe with columns "name", "gdppercapita", "density", "population".
 getdemodata <- function() {
   itprovinces <- getprovincelist()
-  
+
   #' Get GDP/capita data from Eurostat.
   #'
   gdppercapita <- filter(get_eurostat_data(
@@ -140,7 +140,7 @@ getdemodata <- function() {
   ), unit == "EUR_HAB")
   gdppercapita <- select(gdppercapita, c("geo", "values"))
   gdppercapita$geo <- as.character(gdppercapita$geo)
-  
+
   #' Get population density data from Eurostat.
   density <- get_eurostat_data(
     "demo_r_d3dens", # Specific column code for density from Eurostat documentation.
@@ -149,7 +149,7 @@ getdemodata <- function() {
   )
   density <- select(density, c("geo", "values"))
   density$geo <- as.character(density$geo)
-  
+
   #' Get population data from Eurostat.
   population <- filter(get_eurostat_data(
     "demo_r_pjangrp3", # Specific column code for population from Eurostat documentation.
@@ -174,10 +174,10 @@ getdemodata <- function() {
 }
 
 #' Augment Italy DPC data with demographics, and also removes Sardinia provinces
-#' 
+#'
 #' @param dpc Italian per province timeseries data.
 #' @param demodata Dataframe sourced from Eurostat that contains GDP per Capita and Density information on Italian regions.
-#' 
+#'
 #' @return Augmented dataframe that contains both COVID-19 cases timeseries and eurostats.
 augmentDPCdemo <- function(dpc, demodata) {
   dpc <- filter(dpc, lat != 0)  # not regional data
@@ -189,13 +189,14 @@ augmentDPCdemo <- function(dpc, demodata) {
 }
 
 #' Get the closest weather stations to a latitude longitude point.
-#' 
+#'
 #' @param lat Latitude
 #' @param long Longitude
 #' @param n Number of weather stations to get.
-#' 
+#'
 #' @return The codes of the closest weather stations in a list format.
 closestweatherstations <- function(lat, long, n) {
+  print(paste("Querying for", lat, long))
   getMeta(lat = lat, lon = long, n = n)$code
 }
 
@@ -210,11 +211,15 @@ importNOAAM <- memoise(importNOAA)
 #' Augment a single group of data points with weather, using a list of station codes.
 #' @param group A dataframe groupby object on a particular province, which we are pulling weather data for.
[…]

### 2020-04-17 R/DataPipeline.R
comments

@@ -208,6 +208,8 @@ closestweatherstationsM <- memoise(closestweatherstations)
 importNOAAM <- memoise(importNOAA)
 
 #' Augment a single group of data points with weather, using a list of station codes.
+#' Query each station code for air temperature, dewpoint, and relative humidity, taking the closest result that is not NA.
+#' 
 #' @param group A dataframe groupby object on a particular province, which we are pulling weather data for.
 #' @param station A comma separated string of station codes, which are used by importNOAA to query for weather data from those stations.
 #' 
@@ -223,6 +225,8 @@ augmentsingleweather <- function(group, station) {
 }
 
 #' Augment a table with date, lat, and long columns with weather data collected closest to the given time.
+#' For each province, locate the 5 closest weather stations and store their station codes used by NOAA.
+#' Then, query NOAA with augmentsingleweather for the air temperature, dewpoint, and relative humidity at that station.
 #' Warning: This function will be very slow, since it queries the NOAA database often.
 #'
 #' @param dpc Dataframe of DPC per province time series case data. Required columns: "date", "lat", "long".

### 2020-04-17 R/ThompsonModel.R

@@ -1,18 +1,22 @@
-#' Follows the Framework of the Thompson et al paper.
-#'
+#' Follows the Framework of the Thompson et al paper, while using OLS to predict R from climate similar to Wang et al.
+#' Primarily leverages the EpiEstim R package to perform MCMC and estimate R from time series.
+#' Requires some estimate of serial interval, which we simply use values from pervious papers for.
+#' 
 
 library(EpiEstim)
 library(ggplot2)
 
-#' Let's use the Italian Provincial datasets.
-#'
+#' Using Italian Provincial COVID-19 dataset augmented with climate and population statistics.
 dpc_df <- read.csv("data/dpc-augmented.csv")
 dpc_df$date <- as.Date(dpc_df$date, format = "%Y-%m-%d")
 province_names <- unique(dpc_df$province)
 
+#' Prepare an empty dataframe to collect each province's daily R estimates.
 all_provinces_df <- data.frame(matrix(ncol = 7, nrow = 0))
 colnames(all_provinces_df) <- c("date", "province", "mean_R", "gdppercapita", "density", "air_temp","RH")
 
+#' Compute the reproductive rate estimates for each province in our dataset.
+#'
 for (province_name in province_names) {
   region_df <- subset(dpc_df, province == province_name)
   
@@ -43,5 +47,7 @@ for (province_name in province_names) {
 #'
 #' Cutoff on 2020-03-13, since intervention completely changes regime.
 all_provinces_df <- subset(all_provinces_df, date <= "2020-03-13")
+
+#' Perform linear regression on expected R against: GDP per capita, Density, Air Temperature, Relative Humidity.
 fit <- lm(mean_R ~ gdppercapita + density + air_temp + RH, data=all_provinces_df)
 summary(fit)

### 2020-04-17 R/DataPipeline.R
Encapsulate data collection into collectData

@@ -3,6 +3,7 @@ library(dplyr)
 library(worldmet)
 library(memoise)
 library(lubridate)
+library(R.utils)
 
 #' Fetches a raw file from url, and saves it to a filename.
 #' For proper style, filename string should begin with "data/".
@@ -85,10 +86,6 @@ fetch_latest_dpc <- function() {
   names(d)[names(d) == "totale_casi"] <- "total_cases"
   write.csv(d, filename)
 }
-
-fetch_latest_ecdc()
-fetch_latest_dpc()
-fetch_latest_csse()
 
 #' Obtain mapping of health ministry province names to eurostat codes
 #'
@@ -293,18 +290,33 @@ repairtotalcases <- function(dpc) {
   )
 }
 
-#' Code which augments the Italian per province data with GDP, weather, density data.
-#'
-demodata <- getdemodata()
-dpc <- read.csv("data/dpc-covid19-ita-province.csv")
-df <- repairtotalcases(dpc)
-df <- augmentDPCdemo(df, demodata)
-df <- robust(augmentDPCweather, timeout=120)(df)
-write.csv(demodata, "data/demodata.csv")
-write.csv(df, "data/dpc-augmented.csv")
+#' Augments the Italian per province data with GDP, weather, density data.
+#' Produces a file data/dpc-augmented.csv which contains the required data for
+#' this analysis.
+#' @param rewriteall Attempt to refresh all files even if already exist. If
+#'   TRUE, this function may need to be called multiple times due to `worldmet`
+#'   issues.
+collectData <- function(rewriteall = FALSE) {
+  fetch_latest_ecdc()
+  fetch_latest_dpc()
+  fetch_latest_csse()
 
-# test: check that no provinces weren't mapped to demographic data
-stopifnot(nrow(filter(df, is.na(density))) == 0)
+  if (rewriteall | !file.exists("data/demodata.csv")) {
+    demodata <- getdemodata()
+    write.csv(demodata, "data/demodata.csv")
+  }
 
-# test: check no temperature NAs
-stopifnot(nrow(filter(df, is.na(air_temp))) == 0)
+  if (rewriteall | !file.exists("data/dpc-augmented.csv")) {
+    dpc <- read.csv("data/dpc-covid19-ita-province.csv")
+    df <- repairtotalcases(dpc)
+    df <- augmentDPCdemo(df, demodata)
+    df <- robust(augmentDPCweather, timeout=120)(df)
[…]

### 2020-04-17 PaperDraft.Rmd
italian column renaming
write methodology

--- 2020-03-22 ProjectProposal.Rmd
@@
 
-## Background
+## Abstract
+
+## Introduction
+
+[TK: this is retained from Proposal, modify]
 
@@
 
-## Bayesian Framework for Estimation of Time-Varying $R$
+## Methodology
 
-Thompson et al.
-(2019) outline a Bayesian framework for estimation of $R(t)$ using symptomatic case counts, decomposed into local and imported cases.
-The EpiEstim R package implements ideas from this paper.
-A simplified summary of their approach, ignoring the imported cases:
+Thompson et al. (2019) outline a Bayesian framework for estimation of $R(t)$
+using symptomatic case counts, decomposed into local and imported cases. The
+EpiEstim R package implements ideas from this paper. A simplified summary of
+their approach, ignoring the imported cases:
 
-- Estimation of (discrete) distribution of serial interval: Our project does not extend this portion, which estimates the quantities ws that denote the probability density of interval length s under a posterior distribution for the serial interval.
-- Estimation of time-varying $R$: Data is collected containing the number of symptomatic cases observed at time $t$: $I(t)$.
-The observed number of cases at each timestep is assumed to be drawn from a Poisson distribution:
+1. Estimation of (discrete) distribution of serial interval from line-list data
+   with intervals reflecting date of infection.
+2. Estimation of time-varying $R$ from time-bucketed symptomatic case counts.
 
-$$ I(t) \space | \space  I(1), \cdots ,I(t-1),w,R(t) \sim \text{Poisson} \left(R(t) \sum_{s=1}^t I(t-s)w_s \right) $$
+We developed a method to estimate the strength of various seasonal forcings on
+$R$ through MCMC. This involves both substantial deviations from the second part
+of the Thompson et al. model and also less less substantial adjustments for the
+nature of our data. Our method does not extend the first part of the Thompson et
+al. model.
 
-Then Bayesian inference can be used to compute the posterior distribution of $R(t)$ conditional on the data and the serial interval distribution.
-(In the Thompson et al.
-paper, a gamma-distributed prior for $R(t)$ [the conjugate prior for a Poisson likelihood] and a convenient serial interval distribution was chosen so that this posterior is proportional to an analytic expression.)
+To sanity-check the results of our method, we compare its conclusions against a
+point estimate for the seasonal forcings through linear regression with the
+EpiEstim estimates of $R$.
 
-## Extension of Thompson et al. for $R$ subject to forcings
+### Incidence likelihood of the Thompson et al. model
 
-Our proposal is to extend the method of Thompson et al. to multiple locations, each with its own incidence time series $I_1(t),\cdots,I_n(t)$ such that
+In its estimation of time-varying $R$, the Thompson et al. model uses the
+following likelihood function for the number of new infections at time $t$:
 
-$$ R_i(t) = \beta^T x_i(t) + \varepsilon_i(t) $$
+$$ I(t) \space | \space  I(1), \cdots ,I(t-1),\mathbf{w},R(t) \sim \text{Poisson} \left(R(t) \sum_{s=1}^t I(t-s)w_s \right) $$
 
-where $\beta$ is a parameter describing the locally linear effect of each seasonal forcing, $x_i(t)$ is a vector of observed seasonal factors at time $t$ and location $i$, and $\varepsilon_i(t)$ denotes the component of $R_i(t)$ not linearly predicted by seasonal conditions at location $i$.
-and the serial interval distribution $(w_s)_{s \geq 1}$ are assumed to be characteristics of the disease independent of location.
-The observed number of cases at each timestep is
+where
 
-$$ I_i(t) \space | \space I_i(1), \cdots ,I_i(t-1),w,x_i(t), \varepsilon_i(t) \sim \text{Poisson} \left( (\beta^Tx_i(t)+ \varepsilon_i(t)) \sum_{s=1}^t I(t-s) w_s \right) $$
+- $I(t)$ is observed data containing the number of new infections at time $t$
+- $R(t)$ is a parameter for the reproduction number at time $t$
+- $w_s$ denotes the proportion of serial intervals which take on discrete value
+  $s$, i.e. $w_s = \operatorname{P}(\mathrm{SI} = s)$.
 
-Here the parameter of interest is $\beta$ and the $\varepsilon_i(t)$ are nuisance variables.
-We propose to use Markov chain Monte Carlo to compute a joint distribution over all the parameters and marginalize over the $\varepsilon_i(t)$.
-Because of concerns over confounding factors that may affect $R$ and are correlated with temperature and humidity, we plan to add several other features to $x(t)$ to help control for these factors.
-We also need to restrict the analysis to the pre-lockdown regime in each province, because the lockdown is likely to change the magnitude of seasonal forcings.
+Because the Thompson et al. model uses a gamma-distributed prior for $R(t)$ (the
+conjugate prior for a Poisson likelihood) and because they chose a convenient
+family of serial interval distributions, the Bayesian posterior in the Thompson
+et al. model can be described by a closed form analytic expression.
 
-- Time (days)
-- GDP per capita
-- Population density
+### Extension of Thompson et al. for $R$ subject to forcings
 
-## Dataset
+Our Bayesian model extends the method of Thompson et al. to multiple locations,
+each with its own incidence time series $I_1(t),\cdots,I_n(t)$ such that
+
+$$ R_i(t) = \exp\left( \theta(t) + \lambda_i + \mathbf{\beta}^T x_i(t) + \varepsilon_i(t) \right)$$
+
+where:
+
+- $x_i(t)$ is data of observed seasonal factors at time $t$ and location $i$,
+  with each feature standardized to mean $0$ and variance $1$
+- $\beta$ is a parameter describing the locally linear effect of each seasonal
+  forcing
+- $\theta(t)$ is a parameter for logarithm of country-wide $R$ independent of
+  modeled seasonal forcings
+- $\lambda_i$ is a parameter for local variation of $\log R$ independent of
+  modeled seasonal forcings
+- $\varepsilon_i(t)$ is a parameter for remaining fluctuation in $\log R$ not
+  modeled above
+
+Unlike the Thompson et al., our analysis fundamentally estimates the logarithm
+of $R$ instead of $R$ itself. This has some practical advantages: a linear model
+for $R$ could predict negative values for sufficiently extreme climate
+conditions, but negative values for $R$ do not make sense epidemiologically.
+
+We maintain a very similar incidence likelihood function as the Thompson et al.
+model, but impose an additional correction for importations:
+
+$$
+I_i(t) \space |
+\space P_i,x_i(t),I_i(1), \cdots ,I_i(t-1),\mathbf{w},\theta(t),\lambda_i,\mathbf{\beta},\mathbf{\varepsilon}(t),\gamma
+\sim \text{Poisson} \left(\gamma P_i + R_i(t) \sum_{s=1}^t I_i(t-s)w_s \right)
+$$
+
+where in addition to the parameters described (with $R_i(t)$ given by the
+equation above), the following additional data and parameters are introduced:
+
+- $P_i$ is data giving the population of location $i$
+- $\gamma$ is a parameter estimating the nationwide rate of imports per capita
+per day, on which we impose prior $\gamma \sim \operatorname{Gamma}(2, 10^{-5})$
+
+The only substantative change to this likelihood function, beyond the
+decomposition of $R_i(t)$ into the various parameters influencing it, is the
+addition of importation risk. The reason for this adjustment is that the dataset
+we are using does not distinguish between importations and local cases. Thus,
+the Thompson likelihood would be $0$ upon detection of the first case, had we
+not modeled the importations separately.
+
+The most important feature of our model is the parameter $\mathbf{\beta}$ which
+are the coefficients of the linear model of $\log R$. This feature is assumed to
+be global across all locations and times, and thus models inherent properties of
+the contagiousness of the outbreak, as opposed to local deviations in space or
+time.
+
+To avoid the model ascribing deviations in $R$ to the seasonal forcings without
+widespread evidence across locations and times, three other features of our
+model are the parameters $\theta(t)$, $\lambda_i$, and $\varepsilon_i(t)$. The
+first parameter, $\theta(t)$, describes the logarithm country-wide $R$ before
+adjusting for location and climate and local idiosyncratic factors. That is, it
+can be thought of as a baseline $R$ value affected simply by the national
+circumstances at the given time; such as national policies or human behavioral
+changes. This will encapsulate various unmodeled behaviours such as weekends or
+interventions. At time zero, our prior belief is that this $R$ will match the
+consensus of international estimates for $R_0$, which vary widely but tend to
+cluster between $2$ or $3$ [TK: Citation needed?]. Because of the great
+uncertainty, there is significant variance in this assumption, so the prior is
+quite weak (note that the $\mu \pm 2\sigma$ values, when transformed
+exponentially, are below $1$ and above $6$ respectively).
+
+$$
+\theta(0) \sim \operatorname{Normal}(\log 2.5, 0.5^2)
+$$
+
+The remaining days of $\theta$ are modeled using a simple autoregressive model:
+
+$$
+\theta(t+1) \sim \operatorname{Normal}(\theta(t), \theta^\sigma)
+$$
+
+where $\theta^\sigma$ is a parameter describing the spread of daily drifts in
+$\theta$. The prior on $\theta^\sigma$ is weakly informative:
+
+$$
+\theta^\sigma \sim \operatorname{Gamma}(2, 0.01)
+$$
+
+The location-specific deviation from the country average in $\log R$ independent
+of modeled seasonal forcings is encoded in $\lambda_i$. The existence of this
+parameter helps the model explain the large variations of apparent $R$ across
+different locations, some of which appear to be affected much more so than
+others. The remaining parameter $\varepsilon_i(t)$ encapsulates all other
+unexplained deviations across both time and location. All these parameters (and
+$\beta$) are given the following weakly informative prior:
+
+$$
+\lambda_i, \varepsilon_i(t), \beta_k \sim \operatorname{Normal}(0, 0.5)
+$$
+
+Our technique is general and can be used for any epidemic and any combination of
+potential seasonal forcings.
+
+### Dataset
+
+[TK: retained from proposal, need to modify]
 
@@
 
-## Directions for Future Research
+## Results
+
+## Discussion
+
+### Directions for Future Research
+
+[TK: retained from proposal, need to modify]
 
@@
 
-## Citations
+## Appendix
 

### 2020-04-17 R/DataPipeline.R
adding R.utils library to data pipeline

@@ -3,6 +3,7 @@ library(dplyr)
 library(worldmet)
 library(memoise)
 library(lubridate)
+library(R.utils)
 
 #' Fetches a raw file from url, and saves it to a filename.
 #' For proper style, filename string should begin with "data/".

### 2020-04-17 R/ThompsonModel.R
save thompson model as rds file

@@ -51,3 +51,6 @@ all_provinces_df <- subset(all_provinces_df, date <= "2020-03-13")
 #' Perform linear regression on expected R against: GDP per capita, Density, Air Temperature, Relative Humidity.
 fit <- lm(mean_R ~ gdppercapita + density + air_temp + RH, data=all_provinces_df)
 summary(fit)
+
+save(fit, file="data/thompson_fit.rds")
+

### 2020-04-17 PaperDraft.Rmd

@@
 
+As the model is much more complex than the one in the Thompson et al. paper, we
+are unable to derive an analytic expression to compute the posterior. Instead,
+we use `rstan` to run $4$ chains of MCMC for $2000$ iterations each.
+
+### Using OLS and EpiEstim to check results
+
+[TK: @Ben mention the OLS against EpiEstim ]
+
 ### Dataset

### 2020-04-17 R/AnalysisDetails.R
cleanup, package in partially working state

@@ -1,0 +1,29 @@
+#' Description of the seasonal forcing features used in the analysis.
+descr = c("temp", "RH", "GDP per capita", "density")
+
+#' Create labels for dates appropriate for use in a chart, by removing the year.
+#' @param range The date range.
+#' @return A vector of strings to use as labels for a chart.
+removeyear <- function(range) {
+  substr(toString(as.Date(range)), 5, 10)
+}
+
+#' Return a list containing results of the analysis, sufficient for generating plots.
+#' Must be called after `collectData()` and `runMCMC()`.
+#' @return The saved results of the analysis.
+results <- function() {
+  if (!file.exists("data/dpc-augmented.csv")) {
+    throw.error("Could not find augmented DPC data. Did you `collectData()`?")
+  } else if (!file.exists("data/cf_fit.rds") | !file.exists("data/cf_data.rds")) {
+    throw.error("Could not find MCMC fit results. Did you `runMCMC()`?")
+  }
+  dpc <- read.csv("data/dpc-augmented.csv")
+  load("data/cf_fit.rds")
+  load("data/cf_data.rds")
+  list(
+    dpc = dpc,
+    stanfit = cf_fit,
+    standata = cf_data,
+  )
+}
+

### 2020-04-17 R/ClimateForcingModel.R

@@ -3,68 +3,73 @@ library(dplyr)
 library(tidyr)
 library(abind)
 
-# Compile stan model
-cf_mod <- stan_model("ClimateForcingModel.stan")
-save(cf_mod, file="data/cf_mod.rds")
+#' Compile the stan model, run MCMC, and save results.
+#' @param cutoff The cutoff date (default: CUTOFF_DATE).
+function runMCMC(cutoff = "2020-03-14") {
+  cf_mod <- stan_model("ClimateForcingModel.stan")
+  save(cf_mod, file="data/cf_mod.rds")
 
-df <- read.csv("data/dpc-augmented.csv", stringsAsFactors = FALSE)
+  df <- read.csv("data/dpc-augmented.csv", stringsAsFactors = FALSE)
 
-# Stop the analysis on Mar 13
-df <- filter(df, date < "2020-03-14")
+  # Stop the analysis on Mar 13
+  df <- filter(df, date < cutoff)
 
-#' Standardize a single feature.
-#' @param series The series to standardize.
-#' @return A series with mean 0 and standard deviation 1.
-standardize <- function(series) {
-  (series - mean(series)) / sd(series)
+  #' Standardize a single feature.
+  #' @param series The series to standardize.
+  #' @return A series with mean 0 and standard deviation 1.
+  standardize <- function(series) {
+    (series - mean(series)) / sd(series)
+  }
+
+  df <- (
+    df
+    %>% mutate(std_air_temp = standardize(air_temp),
+               std_RH = standardize(RH),
+               std_gdppercapita = standardize(gdppercapita),
+               std_density = standardize(density))
+  )
+
+  # Generate data for climate model.
+  X <- abind(split(
+    select(df, c('std_air_temp', 'std_RH', 'std_gdppercapita', 'std_density')),
+    df$province
+  ), along=3)
+
+  I <- abind(split(df$new_cases, df$province), along=2)
+
+  pop <- abind(df %>% group_by(province) %>% group_map(~ .x$population[1]), along=1)
+
+  #' Discretize the serial interval distribution of 4.7 std. dev 2.9, used in the
+  #' CMMID paper https://epiforecasts.io/covid/methods.html
+  #' @param n The number of discrete dates in the future to compute a serial
+  #'   interval density for.
+  #' @param mean The estimated mean of the serial interval.
+  #' @param sd The estimated standard distribution of the serial interval.
[…]

### 2020-04-17 R/DataPipeline.R

@@ -181,7 +181,7 @@ augmentDPCdemo <- function(dpc, demodata) {
   dpc <- filter(dpc, region != "Sardegna")
 
   df <- left_join(dpc, demodata, by = c("province" = "name"))
-  df <- select(df, c("date", "province", "lat", "long", "gdppercapita", "density", "total_cases", "population"))
+  df <- select(df, c("date", "province", "lat", "long", "gdppercapita", "density", "total_cases", "new_cases", "population"))
   df
 }
 
@@ -207,7 +207,7 @@ importNOAAM <- memoise(importNOAA)
 
 #' Augment a single group of data points with weather, using a list of station codes.
 #' Query each station code for air temperature, dewpoint, and relative humidity, taking the closest result that is not NA.
-#' 
+#'
 #' @param group A dataframe groupby object on a particular province, which we are pulling weather data for.
 #' @param station A comma separated string of station codes, which are used by importNOAA to query for weather data from those stations.
 #'

### 2020-04-17 R/Plots.R

@@ -2,15 +2,6 @@ library(bayesplot)
 library(ggplot2)
 library(gtable)
 
-load("data/cf_fit.rds")
-
-descr = c("temp", "RH", "GDP per capita", "density")
-params <- rstan::extract(cf_fit)
-dates <-
-  c(paste("02-", 24:29, sep = ""),
-    paste("03-0", 1:9, sep = ""),
-    paste("03-", 10:13, sep = ""))
-
 #' Given the Bayesian posterior stanfit object, plot the MCMC confidence for
 #' beta.
 #'
@@ -18,13 +9,14 @@ dates <-
 #' @return A plot of location of standardized beta parameters, with 50%
 #'   confidence interval highlighted.
 plot_beta_confidence <- function(cf_fit) {
+  betadescr <- paste(descr, sep = ", ", collapse = ", ")
   (
     mcmc_areas(
       cf_fit,
       regex_pars = c("beta.+"),
       point_est = "mean"
     ) + ggplot2::labs(title = "Posterior of β, magnitude of climate forcings",
-                      subtitle = "Each standardized; in order: temp., rel. humidity, GDP p.c., density")
+                      subtitle = paste("Each standardized; in order:", betadescr)
   )
 }
 
@@ -148,6 +140,8 @@ plot_epsilon_dispersion <- function(params) {
                   subtitle = "Aggregated across all locations and times",
                   xlab = "difference in log(R)")
 }
+
+# TODO: move these to the Rmarkdown file
 
 plot_beta_confidence(cf_fit)
 plot_lambda_confidence(cf_fit, 1, 34)

### 2020-04-17 PaperDraft.Rmd

@@
 title: "Inference of seasonal forcings on reproduction numbers during epidemics"
-author: "Lu Lu, Ziyao Tian, Fengyang Wang, Yi Xiao Zhang"
+author: "Fengyang Wang, Lu Lu, Ziyao Tian, Yi Xiao Zhang"
 output:
@@
 
+Amidst the current COVID-19 epidemic, the prediction of its future spread and the length of the epidemic becomes increasingly significant.
+We investigate the effects of temperature and humidity on the reproductive rate R of COVID-19, as warmer climates in the summer have historically resulted in diminished spread of influenza.
+Early epidemiology research in COVID-19 have used simple linear models such as ordinary least squares regression to estimate the relationship between R and climate.
+We estimate the reproductive rate of the virus using a novel Bayesian inference model with MCMC sampling to daily case data of various climate regions.
+While results are largely inconclusive, they provide us with a weak posterior update that R decreases with temperature and relative humidity.
+
 ## Introduction
-
-[TK: this is retained from Proposal, modify]
 
@@
 
-- The serial interval: time between onset of symptoms of two adjacent individuals in a transmission chain
-- The reproductive number, $R$: the expected number of new infections a single symptomatic individual will cause
+- The serial interval: time between onset of symptoms of two adjacent individuals in a transmission chain.
+- The reproductive number, $R$: the expected number of new infections a single symptomatic individual will cause.
 
+Due to inadequate public data to compute the serial interval, we use the estimated serial interval of $4.7$ days with standard deviation $2.9$ days obtained from Nishiura et al.
+(2020) [1].
 The estimation of the latter quantity is important for policy, because infectious disease outbreaks will switch from an exponential growth regime to an exponential decay regime when the reproductive number falls below 1.
-The goal of mitigation or suppression policies is therefore to effect change on the reproductive number $R$.
-
+The goal of mitigation or suppression policies is therefore to effect change on the reproductive number R.
 Many infectious diseases are believed to be seasonal, including most respiratory viruses, like the SARS-CoV-2 virus implicated in the 2019–2020 coronavirus pandemic.
 Wang et al.
-(2020) have analysed how the weather of locations within China correlates with local point estimates of the reproductive number $R$.
+(2020) [3] have analysed how the weather of locations within China correlates with local point estimates of the reproductive number R.
 However, because their analysis is frequentist in nature, their estimates of the seasonal forcing do not come with any estimate of uncertainty.
+
+We first produce a baseline model, which predicts the reproductive rate using the EpiEstim package, based on previous work by Thompson et al.
+(2019) [2], and applied to COVID-19 data as in Abbot et al.
+(2020) [4].
+We then apply ordinary least squares regression against climate variables as well as other underlying regional factors, such as population density, GDP per capita, and latitude, similar to the work produced by Wang et al.
+(2020)[3].
+We compare this against our novel Bayesian framework, and evaluate the significance of the explanatory variables in each model.
+
+Since Italy was one of the first countries affected by the epidemic, it has a longer history of case data before intervention and lockdowns took place.
+Furthermore, due to the many provinces of Italy having diverse climates, we have access to a larger range of temperature and humidity.
+Therefore, we sourced the Italian government DPC dataset[5], which contains time series incidence for each province since mid February.
+Using the temperature and humidity data from weather stations sourced from NOAA, we are able to estimate the climate for over $80$ distinct provinces in the Italian COVID-19 dataset.
 
@@
 
-[TK: @Ben mention the OLS against EpiEstim ]
+The baseline model we build estimates the reproductive rate using a parametric SI model, where we specify the mean and standard deviation of the serial interval.
+Because we do not have access to line-list SI data for COVID-19, we are unable to effectively estimate the serial interval from data, thus we take the findings of Nishiura et al.
+(2020), which estimated the mean serial interval to be 4.7 and the standard deviation to be 2.4.
+We use the model detailed in Cori et al.
+(2013) [7] for estimating R from bayesian inference on serial interval and time series data, which can be easily implemented using the EpiEstim package.
 
@@
 
-- (PRE-PRINT) Abbott, S., Hellewell, J., Munday, J.
+### References
+
+[1] Nishiura H, Linton NM, Akhmetzhanov AR. Serial interval of novel coronavirus (2019-nCoV) infections. medRxiv Published Online First: 2020. doi:10.1101/2020.02.03.20019497
+
+[2] R.N.
+Thompson, J.E.
+Stockwin, R.D.
+van Gaalen, J.A.
+Polonsky, Z.N.
+Kamvar, P.A.
+Demarsh, E.
+Dahlqwist, S.
+Li, E.
+Miguel, T.
+Jombart, J.
+Lessler, S.
+Cauchemez, A.
+Cori.
+Improved inference of time-varying reproduction numbers during infectious disease outbreaks, Epidemics, Volume 29, 2019, 100356, ISSN 1755-4365, https://doi.org/10.1016/j.epidem.2019.100356.
+
+[3] Wang, Jingyuan and Tang, Ke and Feng, Kai and Lv, Weifeng, High Temperature and High Humidity Reduce the Transmission of COVID-19 (March 9, 2020). http://dx.doi.org/10.2139/ssrn.3551767
+
+[4] (PREPRINT) Abbott, S., Hellewell, J., Munday, J.
 D., Chun, J.
@@
 CMMID.
-Available at CMMID: [https://cmmid.github.io/topics/covid19/current-patterns-transmission/global-time-varying-transmission.html](https://cmmid.github.io/topics/covid19/current-patterns-transmission/global-time-varying-transmission.html)
+Available at CMMID: https://cmmid.github.io/topics/covid19/current-patterns-transmission/global-time-varying-transmission.html
 
-- Thompson, R.
-N., Stockwin, J.
-E., van Gaalen, R.
-D., Polonsky, J.
-A., Kamvar, Z.
-N., Demarsh, P.
-A., ...
-& Lessler, J.
-(2019).
-Improved inference of time-varying reproduction numbers during infectious disease outbreaks.
-Epidemics, 29, 100356.
+[5] Presidenza del Consiglio dei Ministri - Dipartimento della Protezione Civile, Dati COVID-19 Italia, (2020), GitHub repository, https://github.com/pcm-dpc/COVID-19
 
-- Wang, J., Tang, K., Feng, K., & Lv, W.
-(2020).
-High Temperature and High Humidity Reduce the Transmission of COVID-19.
-Available at SSRN: [https://ssrn.com/abstract=3551767](https://ssrn.com/abstract=3551767) or [http://dx.doi.org/10.2139/ssrn.3551767](http://dx.doi.org/10.2139/ssrn.3551767)
+[6] “Report 13 - Estimating the Number of Infections and the Impact of Non-Pharmaceutical Interventions on COVID-19 in 11 European Countries.” Imperial College London, www.imperial.ac.uk/mrc-global-infectious-disease-analysis/covid-19/report-13-europe-npi-impact/.
+
+[7] Anne Cori, Neil M.
+Ferguson, Christophe Fraser, Simon Cauchemez, A New Framework and Software to Estimate Time-Varying Reproduction Numbers During Epidemics, American Journal of Epidemiology, Volume 178, Issue 9, 1 November 2013, Pages 1505–1512, https://doi.org/10.1093/aje/kwt133

### 2020-04-17 R/AnalysisDetails.R
more toplevel fixes

@@ -14,16 +14,20 @@ removeyear <- function(range) {
 results <- function() {
   if (!file.exists("data/dpc-augmented.csv")) {
     throw.error("Could not find augmented DPC data. Did you `collectData()`?")
-  } else if (!file.exists("data/cf_fit.rds") | !file.exists("data/cf_data.rds")) {
+  } else if (!file.exists("data/cf_fit.rds") | !file.exists("data/cf_data.rds") | !file.exists("data/cf_dates.rds")) {
     throw.error("Could not find MCMC fit results. Did you `runMCMC()`?")
   }
   dpc <- read.csv("data/dpc-augmented.csv")
   load("data/cf_fit.rds")
   load("data/cf_data.rds")
+  load("data/cf_dates.rds")
   list(
     dpc = dpc,
     stanfit = cf_fit,
     standata = cf_data,
+    params = rstan::extract(cf_fit),
+    dates = dates,
+    locations = colnames(cf_data$i)
   )
 }
 

### 2020-04-17 R/ClimateForcingModel.R

@@ -4,7 +4,7 @@ library(tidyr)
 library(abind)
 
 #' Compile the stan model, run MCMC, and save results.
-#' @param cutoff The cutoff date (default: CUTOFF_DATE).
+#' @param cutoff The cutoff date (default: Mar 14).
 function runMCMC(cutoff = "2020-03-14") {
   cf_mod <- stan_model("ClimateForcingModel.stan")
   save(cf_mod, file="data/cf_mod.rds")
@@ -68,6 +68,7 @@ function runMCMC(cutoff = "2020-03-14") {
   # fit stan model
   cf_fit <- sampling(cf_mod, data = cf_data, verbose = TRUE)
 
+  save(unique(df$date), file="data/cf_dates.rds")
   save(cf_fit, file="data/cf_fit.rds")
   save(cf_data, file="data/cf_data.rds")
 

### 2020-04-17 R/DataPipeline.R
drop last 4 days

@@ -278,7 +278,7 @@ robust <- function(f, timeout = 60) {
 }
 
 #' Repair total case data to be monotonically increasing, through taking a rolling maximum.
-#' Drop the last date as it is likely to have no weather data for each location.
+#' Drop the last 4 dates as it is likely to have no weather data for each location.
 #' Add new_cases column which contains the number of new cases each day.
 #'
 #' @param dpc Dataframe of DPC per province timeseries. Requires the columns: "date", "province", "total_cases".
@@ -290,6 +290,9 @@ repairtotalcases <- function(dpc) {
    %>% group_modify(~ arrange(.x, by=date) %>% mutate(total_cases = cummax(total_cases)))
    %>% mutate(new_cases = c(0, diff(total_cases)))
    %>% ungroup
+   %>% filter(date != max(as.character(date)))
+   %>% filter(date != max(as.character(date)))
+   %>% filter(date != max(as.character(date)))
    %>% filter(date != max(as.character(date)))
   )
 }

### 2020-04-17 R/ThompsonModel.R
add csvs

@@ -1,56 +1,57 @@
 #' Follows the Framework of the Thompson et al paper, while using OLS to predict R from climate similar to Wang et al.
 #' Primarily leverages the EpiEstim R package to perform MCMC and estimate R from time series.
 #' Requires some estimate of serial interval, which we simply use values from pervious papers for.
-#' 
+#'
 
 library(EpiEstim)
 library(ggplot2)
 
 #' Using Italian Provincial COVID-19 dataset augmented with climate and population statistics.
-dpc_df <- read.csv("data/dpc-augmented.csv")
-dpc_df$date <- as.Date(dpc_df$date, format = "%Y-%m-%d")
-province_names <- unique(dpc_df$province)
+runThompson <- function() {
+  dpc_df <- read.csv("data/dpc-augmented.csv")
+  dpc_df$date <- as.Date(dpc_df$date, format = "%Y-%m-%d")
+  province_names <- unique(dpc_df$province)
 
-#' Prepare an empty dataframe to collect each province's daily R estimates.
-all_provinces_df <- data.frame(matrix(ncol = 7, nrow = 0))
-colnames(all_provinces_df) <- c("date", "province", "mean_R", "gdppercapita", "density", "air_temp","RH")
+  #' Prepare an empty dataframe to collect each province's daily R estimates.
+  all_provinces_df <- data.frame(matrix(ncol = 7, nrow = 0))
+  colnames(all_provinces_df) <- c("date", "province", "mean_R", "gdppercapita", "density", "air_temp","RH")
 
-#' Compute the reproductive rate estimates for each province in our dataset.
-#'
-for (province_name in province_names) {
-  region_df <- subset(dpc_df, province == province_name)
-  
-  #' Analysis Step 1: Estimate the serial interval
+  #' Compute the reproductive rate estimates for each province in our dataset.
   #'
-  #' We have determined that it is difficult to find the line-list data for COVID-19, and thus will be using
-  #' the mean and standard deviation 
-  est_mean_si = 4.7
-  est_std_si = 2.9
-  
-  #' Analysis Step 2: Estimate the reproduction rate.
+  for (province_name in province_names) {
+    region_df <- subset(dpc_df, province == province_name)
+
+    #' Analysis Step 1: Estimate the serial interval
+    #'
+    #' We have determined that it is difficult to find the line-list data for COVID-19, and thus will be using
+    #' the mean and standard deviation
+    est_mean_si = 4.7
[…]

### 2020-04-17 R/DataPipeline.R
fixed DESCRIPTION that fails devtools run

@@ -284,7 +284,7 @@ robust <- function(f, timeout = 60) {
 #' @param dpc Dataframe of DPC per province timeseries. Requires the columns: "date", "province", "total_cases".
 #'
 #' @return Transformed dataframe with modified "total_cases" column, where the column is now monotonically increasing.
-repairtotalcases <- function(dpc) {
+transform_total_cases <- function(dpc) {
   (dpc
    %>% group_by(province)
    %>% group_modify(~ arrange(.x, by=date) %>% mutate(total_cases = cummax(total_cases)))
@@ -312,7 +312,7 @@ collectData <- function(rewriteall = FALSE) {
 
   if (rewriteall | !file.exists("data/dpc-augmented.csv")) {
     dpc <- read.csv("data/dpc-covid19-ita-province.csv")
-    df <- repairtotalcases(dpc)
+    df <- transform_total_cases(dpc)
     df <- augmentDPCdemo(df, demodata)
     df <- robust(augmentDPCweather, timeout=120)(df)
     write.csv(df, "data/dpc-augmented.csv")

### 2020-04-17 tests/testthat.R

@@ -1,0 +1,4 @@
+library(testthat)
+library(cfc19i)
+
+test_check("cfc19i")

### 2020-04-17 R/AnalysisDetails.R
fix syntax error

@@ -9,25 +9,29 @@ removeyear <- function(range) {
 }
 
 #' Return a list containing results of the analysis, sufficient for generating plots.
-#' Must be called after `collectData()` and `runMCMC()`.
+#' Must be called after `collectData()` and `runMCMC()` and `runThompson()`.
 #' @return The saved results of the analysis.
 results <- function() {
   if (!file.exists("data/dpc-augmented.csv")) {
     throw.error("Could not find augmented DPC data. Did you `collectData()`?")
   } else if (!file.exists("data/cf_fit.rds") | !file.exists("data/cf_data.rds") | !file.exists("data/cf_dates.rds")) {
     throw.error("Could not find MCMC fit results. Did you `runMCMC()`?")
+  } else if (!file.exists("data/thompson_fit.rds")) {
+    throw.error("Could not find Thompson fit results. Did you `runThompson()`?")
   }
   dpc <- read.csv("data/dpc-augmented.csv")
   load("data/cf_fit.rds")
   load("data/cf_data.rds")
   load("data/cf_dates.rds")
+  load("data/thompson_fit.rds")
   list(
     dpc = dpc,
     stanfit = cf_fit,
     standata = cf_data,
     params = rstan::extract(cf_fit),
     dates = dates,
-    locations = colnames(cf_data$i)
+    locations = colnames(cf_data$i),
+    thompsonfit = fit
   )
 }
 

### 2020-04-17 R/ClimateForcingModel.R

@@ -5,7 +5,7 @@ library(abind)
 
 #' Compile the stan model, run MCMC, and save results.
 #' @param cutoff The cutoff date (default: Mar 14).
-function runMCMC(cutoff = "2020-03-14") {
+runMCMC <- function(cutoff = "2020-03-14") {
   cf_mod <- stan_model("ClimateForcingModel.stan")
   save(cf_mod, file="data/cf_mod.rds")
 

### 2020-04-17 R/Plots.R
devtools::install now runs

@@ -16,7 +16,7 @@ plot_beta_confidence <- function(cf_fit) {
       regex_pars = c("beta.+"),
       point_est = "mean"
     ) + ggplot2::labs(title = "Posterior of β, magnitude of climate forcings",
-                      subtitle = paste("Each standardized; in order:", betadescr)
+                      subtitle = paste("Each standardized; in order:", betadescr))
   )
 }
 
@@ -142,15 +142,16 @@ plot_epsilon_dispersion <- function(params) {
 }
 
 # TODO: move these to the Rmarkdown file
-
-plot_beta_confidence(cf_fit)
-plot_lambda_confidence(cf_fit, 1, 34)
-plot_lambda_confidence(cf_fit, 35, 68)
-plot_lambda_confidence(cf_fit, 69, 102)
-plot_independentR(params, cf_data)
-plot_locationR(params, cf_data, 13, "Bergamo")
-plot_locationR(params, cf_data, 55, "Milano")
-plot_locationR(params, cf_data, 79, "Rome")
-plot_locationR(params, cf_data, 96, "Venezia")
-plot_locationR(params, cf_data, 99, "Verona")
-plot_epsilon_dispersion(params)
+generate_plots <- function() {
+  plot_beta_confidence(cf_fit)
+  plot_lambda_confidence(cf_fit, 1, 34)
+  plot_lambda_confidence(cf_fit, 35, 68)
+  plot_lambda_confidence(cf_fit, 69, 102)
+  plot_independentR(params, cf_data)
+  plot_locationR(params, cf_data, 13, "Bergamo")
+  plot_locationR(params, cf_data, 55, "Milano")
+  plot_locationR(params, cf_data, 79, "Rome")
+  plot_locationR(params, cf_data, 96, "Venezia")
+  plot_locationR(params, cf_data, 99, "Verona")
+  plot_epsilon_dispersion(params)
+}

### 2020-04-17 R/ThompsonModel.R

@@ -52,6 +52,5 @@ runThompson <- function() {
   #' Perform linear regression on expected R against: GDP per capita, Density, Air Temperature, Relative Humidity.
   fit <- lm(mean_R ~ gdppercapita + density + air_temp + RH, data=all_provinces_df)
   summary(fit)
-
   save(fit, file="data/thompson_fit.rds")
 }

### 2020-04-17 R/ClimateForcingModel.R
add RDS files

@@ -7,7 +7,8 @@ library(abind)
 #' @param cutoff The cutoff date (default: Mar 14).
 runMCMC <- function(cutoff = "2020-03-14") {
   cf_mod <- stan_model("ClimateForcingModel.stan")
-  save(cf_mod, file="data/cf_mod.rds")
+  options(mc.cores = parallel::detectCores())
+  rstan_options(auto_write = TRUE)
 
   df <- read.csv("data/dpc-augmented.csv", stringsAsFactors = FALSE)
 
@@ -68,7 +69,8 @@ runMCMC <- function(cutoff = "2020-03-14") {
   # fit stan model
   cf_fit <- sampling(cf_mod, data = cf_data, verbose = TRUE)
 
-  save(unique(df$date), file="data/cf_dates.rds")
+  dates <- unique(df$date)
+  save(dates, file="data/cf_dates.rds")
   save(cf_fit, file="data/cf_fit.rds")
   save(cf_data, file="data/cf_data.rds")
 

### 2020-04-17 R/ClimateForcingModel.R
adding stuff to paper

@@ -13,20 +13,14 @@ runMCMC <- function(cutoff = "2020-03-14") {
 
   # Stop the analysis on Mar 13
   df <- filter(df, date < cutoff)
-
-  #' Standardize a single feature.
-  #' @param series The series to standardize.
-  #' @return A series with mean 0 and standard deviation 1.
-  standardize <- function(series) {
-    (series - mean(series)) / sd(series)
-  }
-
+  
+  # Standardize each of the columns we will use.
   df <- (
     df
-    %>% mutate(std_air_temp = standardize(air_temp),
-               std_RH = standardize(RH),
-               std_gdppercapita = standardize(gdppercapita),
-               std_density = standardize(density))
+    %>% mutate(std_air_temp = scale(air_temp),
+               std_RH = scale(RH),
+               std_gdppercapita = scale(gdppercapita),
+               std_density = scale(density))
   )
 
   # Generate data for climate model.

### 2020-04-17 R/ThompsonModel.R

@@ -4,17 +4,29 @@
 #'
 
 library(EpiEstim)
-library(ggplot2)
+library(dplyr)
 
 #' Using Italian Provincial COVID-19 dataset augmented with climate and population statistics.
-runThompson <- function() {
-  dpc_df <- read.csv("data/dpc-augmented.csv")
+#' 
+#' @param filename The path to the augmented data file which we will be using.
+#' @return Fit object, result of linear regression on log(R) against confounding variables.
+runThompson <- function(filename = "data/dpc-augmented.csv") {
+  dpc_df <- read.csv(filename)
   dpc_df$date <- as.Date(dpc_df$date, format = "%Y-%m-%d")
   province_names <- unique(dpc_df$province)
 
   #' Prepare an empty dataframe to collect each province's daily R estimates.
   all_provinces_df <- data.frame(matrix(ncol = 7, nrow = 0))
   colnames(all_provinces_df) <- c("date", "province", "mean_R", "gdppercapita", "density", "air_temp","RH")
+  
+  #' Standardize each of the variables.
+  dpc_df <- (
+    dpc_df
+    %>% mutate(air_temp = scale(air_temp),
+               RH = scale(RH),
+               gdppercapita = scale(gdppercapita),
+               density = scale(density))
+  )
 
   #' Compute the reproductive rate estimates for each province in our dataset.
   #'
@@ -24,14 +36,14 @@ runThompson <- function() {
     #' Analysis Step 1: Estimate the serial interval
     #'
     #' We have determined that it is difficult to find the line-list data for COVID-19, and thus will be using
-    #' the mean and standard deviation
+    #' the mean and standard deviation used by Abbott et al (2020).
     est_mean_si = 4.7
     est_std_si = 2.9
 
     #' Analysis Step 2: Estimate the reproduction rate.
     #'
     res_parametric_si <- estimate_R(region_df$new_cases,
-          method="parametric_si",
+          method = "parametric_si",
           config = make_config(list(
             mean_si = est_mean_si,
             std_si = est_std_si)))
@@ -49,8 +61,12 @@ runThompson <- function() {
   #' Cutoff on 2020-03-13, since intervention completely changes regime.
   all_provinces_df <- subset(all_provinces_df, date <= "2020-03-13")
 
+  #' Take logarithm of mean_R.
+  all_provinces_df$mean_R <- log(all_provinces_df$mean_R)
+  
[…]

### 2020-04-17 tests/testthat/test-Plots.R
add a test

@@ -1,3 +1,7 @@
-test_that("multiplication works", {
-  expect_equal(2 * 2, 4)
+test_that("80% confidence interval works", {
+  data <- 1:100
+  results <- mean_cl_quantile(data, na.rm = TRUE)
+  expect_equal(results$y, 50.5)
+  expect_equal(results$ymin, 10.9)
+  expect_equal(results$ymax, 90.1)
 })

### 2020-04-17 tests/testthat/test-Plots.R
another plots test

@@ -5,3 +5,12 @@ test_that("80% confidence interval works", {
   expect_equal(results$ymin, 10.9)
   expect_equal(results$ymax, 90.1)
 })
+
+test_that("remove year of date range works on various types", {
+  range <- c("2019-07-02", "2019-07-03", "2030-01-01")
+  ry1 <- removeyear(range)
+  ry2 <- removeyear(as.POSIXct(range))
+  expect_equal(ry1, c("07-02", "07-03", "01-01"))
+  expect_equal(ry2, c("07-02", "07-03", "01-01"))
+})
+

### 2020-04-17 tests/testthat/test-DataPipeline.R
another test

@@ -1,0 +1,12 @@
+test_that("transform total cases works", {
+  province <- c(rep("A", 9), rep("B", 9))
+  dateraw <- c(paste("2020-01-0", 1:9, sep=""))
+  date <- c(dateraw, dateraw)
+  total_cases <- c(0, 2, 3, 5, 9, 12, 15, 19, 20,
+                   3, 5, 3, 9, 2, 5, 100, 120, 140)
+  df <- data.frame(province, date, total_cases)
+  df <- transform_total_cases(df)
+  expect_equal(as.character(df$date), as.character(c(dateraw[1:5], dateraw[1:5])))
+  expect_equal(df$total_cases, c(0, 2, 3, 5, 9, 3, 5, 5, 9, 9))
+  expect_equal(df$new_cases, c(0, 2, 1, 2, 4, 0, 2, 0, 4, 0))
+})

### 2020-04-17 PaperDraft.Rmd
adding stuff to paper
jojo graphs

@@
 author: "Fengyang Wang, Lu Lu, Ziyao Tian, Yi Xiao Zhang"
+header-includes:
+    - \usepackage{setspace}\onehalfspacing
 output:
@@
   html_notebook: default
+  rmarkdown::pdf_document:
+    fig_caption: yes        
+    includes:  
+      in_header: preamble.tex
 ---
@@
 - The `worldmet` R package enables access to temperature and humidity data, sourced from NOAA.
+-
 
 ## Results
+[What methods/models are best]
+[What condition(s) does each method/model work well/poorly]
+[Conclusion about dataset]
+
+Applying the methods depicted from the methodology section, we obtain the following results.
+
+\begin{center}
+\includegraphics[width=0.5\textwidth]{plots/plot1.png}
+\end{center}
+
+[TK: Beta values]
+Fitting our Bayesian model, which extends the method of Thompson et.
+al, values of beta were estimated for temperature, relative humidity, Gross Domestic Product (GDP) and population density, respectively.
+Thus, our model estimates $\beta^T = [,,,,]$ in
+$$ R_i(t) = \exp\left( \theta(t) + \lambda_i + \mathbf{\beta}^T x_i(t) + \varepsilon_i(t) \right)$$
+Note that each feature's data is standardized to mean 0 and variance 1.
+
+The model indicates that higher temperatures are observed with lower $R_0$ values, and higher GDP is observed with higher $R_0$ values, while relative humidity and population density have $R_0$ close to zero.
+
+### Posterior of Lambda, location-dependent adjustment to log(R)
+
+\begin{center}
+\includegraphics[width=.49\textwidth]{plots/plot2.png}
+\includegraphics[width=.49\textwidth]{plots/plot3.png}
+\includegraphics[width=.49\textwidth]{plots/plot4.png}
+\end{center}
+
+- $\lambda$ is an adjustment to the model based on the location you're in. 
+### Climate-Independent Basic R, by Date
+
+\begin{center}
+\includegraphics[width=.49\textwidth]{plots/plot5.png}
+\end{center}
+- R_0 is high in early phases, with an early spike on February 27, potentially due to increased in testing, which captures cases before.
+- Climate-independent means ignores local climates, and produces an aggregate result.
+
+### Estimated R in Cities by Date
+
+\begin{center}
+\includegraphics[width=.49\textwidth]{plots/plot6.png}
+\includegraphics[width=.49\textwidth]{plots/plot7.png}
+\includegraphics[width=.49\textwidth]{plots/plot8.png}
+\includegraphics[width=.49\textwidth]{plots/plot9.png}
+\includegraphics[width=.49\textwidth]{plots/plot10.png}
+\end{center}
+
+- local, no longer climate independent.
+
+### Epsilon, Idiosyncratic adjustment to log(R)
+\begin{center}
+\includegraphics[width=.49\textwidth]{plots/plot11.png}
+\end{center}
+
+- You want to minimize the value episilon.
 
@@
 ### Directions for Future Research
+
+As COVID-19 is also an ongoing epidemic, more data surfaces and updates on a continuous basis.
+Future research directions consist of updating the dataset by extending the time range to include newer data and including additional countries and cities - such as by looking at the climate and infection data in American and Canadian cities.
+In particular, by looking at the results of other cities and making a comparison with the results in Italy, we can observe if the linear assumption between seasonal factors and R holds true across different climate regions.
+If not, we can try fitting while assuming a nonlinear relationship between R and seasonal factors.
+Additional future research includes looking into statistical simulation models which consider intercity travel and geographic proximity.
+Using the results from the simulation models, we can help further isolate the direct effects of temperature and humidity on the reproduction rate of COVID-19.
 

### 2020-04-17 PaperDraft.Rmd
some more shitty changes to the paper

--- 2020-04-17 PaperDraft.Rmd
@@
 
-[TK: retained from proposal, need to modify]
+One particularly counterintuitive result in
 
-- Looking at the climate and infection data in American and Canadian cities as the situation develops further.
+With our `cfc19i` library, it is
+
+Looking at the climate and infection data in American and Canadian cities as the situation develops further.
 In particular, compare this with the results in Italy, and observe if the linear assumption between seasonal factors and $R$ holds true across different climate regions.
 If not, try fitting while assuming a nonlinear relationship between $R$ and seasonal factors.
-- Look into statistical simulation models which consider intercity travel and geographic proximity.
+
+Look into statistical simulation models which consider intercity travel and geographic proximity.
 

### 2020-04-17 tests/testthat/test-DataPipeline.R
fix test-DataPipeline

@@ -1,3 +1,5 @@
+library(dplyr)
+
 test_that("transform total cases works", {
   province <- c(rep("A", 9), rep("B", 9))
   dateraw <- c(paste("2020-01-0", 1:9, sep=""))

### 2020-04-18 PaperDraft.Rmd
fix test-DataPipeline

--- 2020-04-17 PaperDraft.Rmd
@@
 
-As COVID-19 is also an ongoing epidemic, more data surfaces and updates on a continuous basis.
-Future research directions consist of updating the dataset by extending the time range to include newer data and including additional countries and cities - such as by looking at the climate and infection data in American and Canadian cities.
-In particular, by looking at the results of other cities and making a comparison with the results in Italy, we can observe if the linear assumption between seasonal factors and R holds true across different climate regions.
-If not, we can try fitting while assuming a nonlinear relationship between R and seasonal factors.
-Additional future research includes looking into statistical simulation models which consider intercity travel and geographic proximity.
-Using the results from the simulation models, we can help further isolate the direct effects of temperature and humidity on the reproduction rate of COVID-19.
+One particularly counterintuitive result in de
 
-[TK: retained from proposal, need to modify]
+With our `cfc19i` library, it is
 
-- Looking at the climate and infection data in American and Canadian cities as the situation develops further.
+Looking at the climate and infection data in American and Canadian cities as the situation develops further.
 In particular, compare this with the results in Italy, and observe if the linear assumption between seasonal factors and $R$ holds true across different climate regions.
 If not, try fitting while assuming a nonlinear relationship between $R$ and seasonal factors.
-- Look into statistical simulation models which consider intercity travel and geographic proximity.
+
+Look into statistical simulation models which consider intercity travel and geographic proximity.
 

### 2020-04-18 R/ClimateForcingModel.R
many package fixes

@@ -3,6 +3,20 @@ library(dplyr)
 library(tidyr)
 library(abind)
 
+#' Discretize the serial interval distribution of 4.7 std. dev 2.9, used in the
+#' CMMID paper https://epiforecasts.io/covid/methods.html
+#' @param n The number of discrete dates in the future to compute a serial
+#'   interval density for.
+#' @param mean The estimated mean of the serial interval.
+#' @param sd The estimated standard distribution of the serial interval.
+#' @return A vector whose index corresponds to the number of days in the future,
+#'   and whose value corresponds to the density of the serial interval
+#'   distribution.
+getSIdist <- function(n, mean = 4.7, sd = 2.9) {
+  w <- dnorm(1:n, mean, sd)
+  w / sum(w)
+}
+
 #' Compile the stan model, run MCMC, and save results.
 #' @param cutoff The cutoff date (default: Mar 14).
 runMCMC <- function(cutoff = "2020-03-14") {
@@ -14,7 +28,7 @@ runMCMC <- function(cutoff = "2020-03-14") {
 
   # Stop the analysis on Mar 13
   df <- filter(df, date < cutoff)
-  
+
   # Standardize each of the columns we will use.
   df <- (
     df
@@ -34,19 +48,6 @@ runMCMC <- function(cutoff = "2020-03-14") {
 
   pop <- abind(df %>% group_by(province) %>% group_map(~ .x$population[1]), along=1)
 
-  #' Discretize the serial interval distribution of 4.7 std. dev 2.9, used in the
-  #' CMMID paper https://epiforecasts.io/covid/methods.html
-  #' @param n The number of discrete dates in the future to compute a serial
-  #'   interval density for.
-  #' @param mean The estimated mean of the serial interval.
-  #' @param sd The estimated standard distribution of the serial interval.
-  #' @return A vector whose index corresponds to the number of days in the future,
-  #'   and whose value corresponds to the density of the serial interval
-  #'   distribution.
-  getSIdist <- function(n, mean = 4.7, sd = 2.9) {
-    w <- dnorm(1:n, mean, sd)
-    w / sum(w)
-  }
   w <- getSIdist(nrow(I))
 
   # stan data

### 2020-04-18 R/Plots.R

@@ -49,6 +49,8 @@ plot_beta_confidence <- function(res) {
 #' lambda
 #'
 #' @param res Output of `results()` containing results of the analysis.
+#' @param start First location to display.
+#' @param last Last location to display.
 #' @return A plot of location of standardized beta parameters, with 50%
 #'   confidence interval highlighted.
 plot_lambda_confidence <- function(res, start, end) {
@@ -107,7 +109,7 @@ plot_independentR <- function(res) {
 #' @param res Samples of posterior distribution of parameters.
 #' @param loc Index of the location to examine.
 #' @return A ggplot2 time series of estimated R with 80% confidence.
-plot_locationR <- function(params, loc) {
+plot_locationR <- function(res, loc) {
   thetadf <- data.frame(res$params$theta +
                           res$params$lambda[loc] +
                           res$params$epsilon[, loc, ] +

### 2020-04-18 tests/testthat/test-DataPipeline.R

@@ -1,12 +1,34 @@
-test_that("transform total cases works", {
+library(dplyr)
+
+#' Create a mock DPC-like dataframe for testing data pipeline.
+createmockdpc <- function() {
   province <- c(rep("A", 9), rep("B", 9))
+  region <- rep("C", 18)
+  lat <- c(rep(1.0, 9), rep(2.0, 9))
+  long <- c(rep(1.0, 9), rep(2.0, 9))
   dateraw <- c(paste("2020-01-0", 1:9, sep=""))
   date <- c(dateraw, dateraw)
   total_cases <- c(0, 2, 3, 5, 9, 12, 15, 19, 20,
                    3, 5, 3, 9, 2, 5, 100, 120, 140)
-  df <- data.frame(province, date, total_cases)
+  df <- data.frame(province, region, date, total_cases, lat, long)
+}
+
+test_that("transform total cases works", {
+  df <- createmockdpc()
   df <- transform_total_cases(df)
-  expect_equal(as.character(df$date), as.character(c(dateraw[1:5], dateraw[1:5])))
+  expect_equal(nrow(df), 10)
   expect_equal(df$total_cases, c(0, 2, 3, 5, 9, 3, 5, 5, 9, 9))
   expect_equal(df$new_cases, c(0, 2, 1, 2, 4, 0, 2, 0, 4, 0))
+})
+
+test_that("augment demographic data works", {
+  df <- createmockdpc()
+  df <- transform_total_cases(df)
+  density <- gdppercapita <- c(1.0, 2.0)
+  population <- c(100, 50)
+  name <- c("A", "B")
+  demodata <- data.frame(density, gdppercapita, name, population)
+  df <- augmentDPCdemo(df, demodata)
+  expect_equal(nrow(df), 10)
+  expect_equal(nrow(filter(df, population == 100)), 5)
 })

### 2020-04-18 PaperDraft.Rmd

@@
 
-### Dataset
-
-[TK: retained from proposal, need to modify]
-
-- The Italian government dataset at https://github.com/pcm-dpc/COVID-19 contains the disease incidence for each province.
-This data contains date of confirmation instead of date of onset of symptoms, and so some procedure may be necessary to correct for this.
-Abbott et al.
-(2020) outlines a technique to correct for reporting delays using an exponential distribution of reporting delay fit to line list data; we may use the same distribution.
-- The `worldmet` R package enables access to temperature and humidity data, sourced from NOAA.
--
+For each province, we take a 7 day rolling sum on the number of new cases each day, and feed the data into the parametric SI model.
 
@@
 
-One particularly counterintuitive result in de
-
 With our `cfc19i` library, it is

### 2020-04-18 R/ClimateForcingModel.R
One more test, move STAN model to top-level to compile it earlier

@@ -3,6 +3,13 @@ library(dplyr)
 library(tidyr)
 library(abind)
 
+options(mc.cores = parallel::detectCores())
+rstan_options(auto_write = TRUE)
+
+#' Compiled STAN model of the seasonal forcing model described in accompanying
+#' report.
+cf_mod <- stan_model("ClimateForcingModel.stan")
+
 #' Discretize the serial interval distribution of 4.7 std. dev 2.9, used in the
 #' CMMID paper https://epiforecasts.io/covid/methods.html
 #' @param n The number of discrete dates in the future to compute a serial
@@ -20,10 +27,6 @@ getSIdist <- function(n, mean = 4.7, sd = 2.9) {
 #' Compile the stan model, run MCMC, and save results.
 #' @param cutoff The cutoff date (default: Mar 14).
 runMCMC <- function(cutoff = "2020-03-14") {
-  cf_mod <- stan_model("ClimateForcingModel.stan")
-  options(mc.cores = parallel::detectCores())
-  rstan_options(auto_write = TRUE)
-
   df <- read.csv("data/dpc-augmented.csv", stringsAsFactors = FALSE)
 
   # Stop the analysis on Mar 13

### 2020-04-18 PaperDraft.Rmd
results

@@
 Because we do not have access to line-list SI data for COVID-19, we are unable to effectively estimate the serial interval from data, thus we take the findings of Nishiura et al.
-(2020), which estimated the mean serial interval to be 4.7 and the standard deviation to be 2.4.
-We use the model detailed in Cori et al.
-(2013) [7] for estimating R from bayesian inference on serial interval and time series data, which can be easily implemented using the EpiEstim package.
+(2020) [1], which estimated the mean serial interval to be 4.7 and the standard deviation to be 2.4.
+We use the model detailed in Wallinga and Teunis (2004) [8] for estimating R from bayesian inference on a fixed serial interval distribution and time series data, which can be easily implemented using the EpiEstim package produced by Cori et al.
+(2013) [7].
 
-For each province, we take a 7 day rolling sum on the number of new cases each day, and feed the data into the parametric SI model.
+For each province, we take a 7 day rolling sum on the number of new cases each day, and feed the data into the parametric SI model, which is given distribution $w(\tau)$ From there, the model uses a likelihood based approach to estiamte R at each discrete timestep, computing
 
+$$
+p_{ij} = \frac{w(t_i - t_j)}{\sum_{i \neq k} }
+$$
 ## Results
@@
 Ferguson, Christophe Fraser, Simon Cauchemez, A New Framework and Software to Estimate Time-Varying Reproduction Numbers During Epidemics, American Journal of Epidemiology, Volume 178, Issue 9, 1 November 2013, Pages 1505–1512, https://doi.org/10.1093/aje/kwt133
+
+[8] Jacco Wallinga, Peter Teunis, Different Epidemic Curves for Severe Acute Respiratory Syndrome Reveal Similar Impacts of Control Measures, American Journal of Epidemiology, Volume 160, Issue 6, 15 September 2004, Pages 509–516, https://doi.org/10.1093/aje/kwh255

### 2020-04-18 R/ThompsonModel.R

@@ -1,4 +1,5 @@
-#' Follows the Framework of the Thompson et al paper, while using OLS to predict R from climate similar to Wang et al.
+#' Follows the framework of the Thompson et al. (2019) paper, while using OLS to predict R from climate similar to Wang et al.
+#' EpiEstim's parametric_si mode is actually based on work done by Wallinga and Teunis (2004).
 #' Primarily leverages the EpiEstim R package to perform MCMC and estimate R from time series.
 #' Requires some estimate of serial interval, which we simply use values from pervious papers for.
 #'

### 2020-04-18 PaperDraft.Rmd
update

@@
 $$
+
 ## Results
+
 [What methods/models are best]
@@
 [Conclusion about dataset]
+
+### Posterior Estimates
 
@@
 [TK: Beta values]
-Fitting our Bayesian model, which extends the method of Thompson et.
-al, values of beta were estimated for temperature, relative humidity, Gross Domestic Product (GDP) and population density, respectively.
-Thus, our model estimates $\beta^T = [,,,,]$ in
+Fitting our Bayesian model, which extends the method of Thompson et. al, values
+of beta were estimated for temperature, relative humidity, Gross Domestic
+Product (GDP) and population density, respectively. The mean estimate of $\beta^T$ is
+
+$$ \beta^T = [-0.23198507 -0.05391226  0.26336295 -0.03586852] $$
+
+in the overall model for $R$
+
 $$ R_i(t) = \exp\left( \theta(t) + \lambda_i + \mathbf{\beta}^T x_i(t) + \varepsilon_i(t) \right)$$
@@
 
-The model indicates that higher temperatures are observed with lower $R_0$ values, and higher GDP is observed with higher $R_0$ values, while relative humidity and population density have $R_0$ close to zero.
+The model indicates that higher temperatures are observed with lower $R$ values,
+and higher GDP is observed with higher $R$ values, while relative humidity and
+population density have close to no effect on $R$. Because of the confidence
+intervals, we are fairly confident of the directional correctness of this
+relationship for temperature and for GDP per capita.
 
-### Posterior of Lambda, location-dependent adjustment to log(R)
-
-\begin{center}
-\includegraphics[width=.49\textwidth]{plots/plot2.png}
-\includegraphics[width=.49\textwidth]{plots/plot3.png}
-\includegraphics[width=.49\textwidth]{plots/plot4.png}
-\end{center}
-
-- $\lambda$ is an adjustment to the model based on the location you're in.
 ### Climate-Independent Basic R, by Date
@@
 \end{center}
-- R_0 is high in early phases, with an early spike on February 27, potentially due to increased in testing, which captures cases before.
-- Climate-independent means ignores local climates, and produces an aggregate result.
+
+$R$ is high in early phases, with an early spike on February 27, potentially due
+to increases in testing, which means the data (and hence the model) has
+underestimated the number of infectious cases compared to the ground truth.
+This result is aggregate across the entire nation and is net of any contribution
+to national weather changes.
 
@@
 
-- local, no longer climate independent.
+These plots show the estimated $R$ values in each province computed through the
+samples from the MCMC. An important observation is that the confidence intervals
+for many provinces are very wide, because those provinces have relatively little
+data. However, one strength of our model is that no weighting on results from
+each province needs to be manually done. The data from larger provinces is
+automatically more important because the Poisson likelihood function returns
+smaller results as the rate parameter increases.
 
-### Epsilon, Idiosyncratic adjustment to log(R)
-\begin{center}
-\includegraphics[width=.49\textwidth]{plots/plot11.png}
-\end{center}
-
-- You want to minimize the value episilon.
+We have also analyzed the $\lambda$ and $\varepsilon$ parameters of the model,
+but because these are not as important, see the Appendix for plots and analysis.
 
@@
 ## Appendix
+
+### Posterior of $\lambda$, location-dependent adjustment to $log(R)$
+
+\begin{center}
+\includegraphics[width=.49\textwidth]{plots/plot2.png}
+\includegraphics[width=.49\textwidth]{plots/plot3.png}
+\includegraphics[width=.49\textwidth]{plots/plot4.png}
+\end{center}
+
+$\lambda$ is an adjustment to the model based on the location. What we observe
+from the distribution of $\lambda$ is that some locations have much more
+certainty in how they deviate from the mean than others. This can be explained
+by the unequal distribution of cases across provinces; provinces with more cases
+lend themselves to more accurate estimation of $R$.
+
+### Epsilon, idiosyncratic adjustment to log(R)
+
+\begin{center}
+\includegraphics[width=.49\textwidth]{plots/plot11.png}
+\end{center}
+
+- You want to minimize the value episilon.
 

### 2020-04-18 PaperDraft.Rmd
moved some sections to appendix

--- 2020-04-18 PaperDraft.Rmd
@@
 
-For each province, we take a 7 day rolling sum on the number of new cases each day, and feed the data into the parametric SI model, which is given distribution $w(\tau)$ From there, the model uses a likelihood based approach to estiamte R at each discrete timestep, computing
+For each province, we take a 7 day rolling sum on the number of new cases each day, and feed the data into the parametric SI model, which is given distribution $w(\tau)$ From there, the model uses a likelihood based approach to estiamte R at each discrete timestep, computing the likelihood that a person at timestep $i$ is infected by a person infected at timestep $j$, and summing the likelihood for each $j$ across all its infectees to compute the $R$ value at timestep $j$.
+This method is outlined in detail in Wallinga and Teunis (2004)[8].
 
-$$
-p_{ij} = \frac{w(t_i - t_j)}{\sum_{i \neq k} }
-$$
 ## Results
@@
 
-[TK: Beta values]
+[TK: Beta Values]
 Fitting our Bayesian model, which extends the method of Thompson et.
@@
 
-### Posterior of Lambda, location-dependent adjustment to log(R)
-
-\begin{center}
-\includegraphics[width=.49\textwidth]{plots/plot2.png}
-\includegraphics[width=.49\textwidth]{plots/plot3.png}
-\includegraphics[width=.49\textwidth]{plots/plot4.png}
-\end{center}
-
-- $\lambda$ is an adjustment to the model based on the location you're in.
 ### Climate-Independent Basic R, by Date
@@
 \end{center}
-- R_0 is high in early phases, with an early spike on February 27, potentially due to increased in testing, which captures cases before.
-- Climate-independent means ignores local climates, and produces an aggregate result.
 
-### Estimated R in Cities by Date
+First, we observe that our computed value of $R_0$ in Italy matches what We observe that R_0 spikes in , with an early spike on February 27, potentially due to increased in testing, which captures cases before.
+
+### Estimated R in Various Provinces
 
@@
 
-- local, no longer climate independent.
-
-### Epsilon, Idiosyncratic adjustment to log(R)
-\begin{center}
-\includegraphics[width=.49\textwidth]{plots/plot11.png}
-\end{center}
-
-- You want to minimize the value episilon.
+Observe that in many cities, our estimated $R_0$ values are extremely high
 
@@
 ## Appendix
+
+### Posterior of Lambda, location-dependent adjustment to log(R)
+
+\begin{center}
+\includegraphics[width=.49\textwidth]{plots/plot2.png}
+\includegraphics[width=.49\textwidth]{plots/plot3.png}
+\includegraphics[width=.49\textwidth]{plots/plot4.png}
+\end{center}
+
+### Epsilon, Idiosyncratic adjustment to log(R)
+\begin{center}
+\includegraphics[width=.49\textwidth]{plots/plot11.png}
+\end{center}
 

### 2020-04-18 PaperDraft.Rmd

--- 2020-04-18 PaperDraft.Rmd
@@
 
-For each province, we take a 7 day rolling sum on the number of new cases each day, and feed the data into the parametric SI model, which is given distribution $w(\tau)$ From there, the model uses a likelihood based approach to estiamte R at each discrete timestep, computing
-
-$$
-p_{ij} = \frac{w(t_i - t_j)}{\sum_{i \neq k} }
-$$
+For each province, we take a 7 day rolling sum on the number of new cases each day, and feed the data into the parametric SI model, which is given distribution $w(\tau)$ From there, the model uses a likelihood based approach to estiamte R at each discrete timestep, computing the likelihood that a person at timestep $i$ is infected by a person infected at timestep $j$, and summing the likelihood for each $j$ across all its infectees to compute the $R$ value at timestep $j$.
+This method is outlined in detail in Wallinga and Teunis (2004)[8].
 
@@
 
-[TK: Beta values]
 Fitting our Bayesian model, which extends the method of Thompson et. al, values
@@
 
-in the overall model for $R$
+in the overall model for $R$:
 
 $$ R_i(t) = \exp\left( \theta(t) + \lambda_i + \mathbf{\beta}^T x_i(t) + \varepsilon_i(t) \right)$$
+
 Note that each feature's data is standardized to mean 0 and variance 1.
@@
 
-### Climate-Independent Basic R, by Date
+### Estimates of $\theta$, transformed to climate-independent $R$ by date
 
@@
 
-### Estimated R in Cities by Date
+### Estimated R in Various Provinces
 
@@
 \includegraphics[width=.49\textwidth]{plots/plot9.png}
-\includegraphics[width=.49\textwidth]{plots/plot10.png}
 \end{center}
@@
 but because these are not as important, see the Appendix for plots and analysis.
+
+In many cities, our estimated $R_0$ values are extremely high. This is because
+those cities reported a dramatic increase in cases, and our estimated $R_0$
+values are indeed consistent with the data. However, we will discuss in the
+interpretation section the limits of our approach, and why we believe this may
+not match the ground truth.
 
@@
 
-- You want to minimize the value episilon.
+This histogram shows all samples of all entries in the $\varepsilon$ matrix.
+Generally, we prefer a lower average spread of $\varepsilon$, because it means
+that more of the variation in reproduction number can be predicted from the
+seasonal factors, location, and time alone. Unfortunately, we found the observed
+spread of $\varepsilon$ is quite high, indicating that there is substantial
+variance in reproductive numbers inferred from the data which cannot be
+explained by modeled features.
 

### 2020-04-18 PaperDraft.Rmd
add on discussion

@@
 For each province, we take a 7 day rolling sum on the number of new cases each day, and feed the data into the parametric SI model, which is given distribution $w(\tau)$ From there, the model uses a likelihood based approach to estiamte R at each discrete timestep, computing the likelihood that a person at timestep $i$ is infected by a person infected at timestep $j$, and summing the likelihood for each $j$ across all its infectees to compute the $R$ value at timestep $j$.
+Thus, for each province, we produce a daily time series estimate for $R$ on each day.
 This method is outlined in detail in Wallinga and Teunis (2004)[8].
 
+For each province, day pair, we have $\bar{x} = (x_1, x_2, x_3, x_4), y$ where:
+
+- $x_1$: standardized GDP per capita
+- $x_2$: standardized population density
+- $x_3$: standardized air temperature
+- $x_4$: standardized relative humidity
+- $y$: logarithm of estimated R
+
+We then perform ordinary least squares regression to find $\bar{\beta}$:
+$$
+y = \beta_0 + \beta_1x_1 + \beta_2x_2 + \beta_3 x_3 + \beta_4 x_4 + \varepsilon
+$$
+
 ## Results
-
-[What methods/models are best]
-[What condition(s) does each method/model work well/poorly]
-[Conclusion about dataset]
 
@@
 
+### Interpretation
+
+While the overall results are not surprising, some caution is important in the
+interpretation of these results. There are at least three sources of uncertainty
+not captured in the analysis:
+
+- Uncertainty about the serial interval distribution (our method assumes the
+  serial interval distribution is known)
+- Uncertainty about timeliness of data (our model assumes cases are reported on
+  the date of infection; in reality there is a significant delay, which the model
+  does not attempt to adjust for)
+- Uncertainty of the completeness of the data (our method assumes all cases are
+  reported, which may not be true, as the high spikes in estimated $R$ are
+  unlikely to be ground truth; a more likely explanation is that early cases went
+  undetected for a long time)
+- Uncertainty about the accuracy and currency of data (our model assumes that
+  temperature, relative humidity, GDP per capita, and population data are local
+  and current, but in reality the demographic data is from 2017 and the weather
+  data is measured by the closest station that may be in some cases not even
+  located in the province)
+
+Hence, it is likely that the uncertainty in posterior estimations of $\beta$ by
+the model is understated. While the model provides some evidence that
+temperature is effective at reducing spread, we do not believe this evidence to
+be as strong as claimed.
+
 ## Discussion
+Our model suggests that there is some weak evidence that with increasing temperature, the reproductive number R decreases.
+The effect of relative humidity on the reproductive number is much less conclusive.
+As we only used data from Italy in our model building, there is a limited range of weather information than we would like to have.
 

### 2020-04-18 PaperDraft.Rmd
Uncertainty analysis

@@
 For each province, we take a 7 day rolling sum on the number of new cases each day, and feed the data into the parametric SI model, which is given distribution $w(\tau)$ From there, the model uses a likelihood based approach to estiamte R at each discrete timestep, computing the likelihood that a person at timestep $i$ is infected by a person infected at timestep $j$, and summing the likelihood for each $j$ across all its infectees to compute the $R$ value at timestep $j$.
-Thus, for each province, we produce a daily time series estimate for $R$ on each day.
 This method is outlined in detail in Wallinga and Teunis (2004)[8].
 
-For each province, day pair, we have $\bar{x} = (x_1, x_2, x_3, x_4), y$ where:
+## Results
 
-- $x_1$: standardized GDP per capita
-- $x_2$: standardized population density
-- $x_3$: standardized air temperature
-- $x_4$: standardized relative humidity
-- $y$: logarithm of estimated R
-
-We then perform ordinary least squares regression to find $\bar{\beta}$:
-$$
-y = \beta_0 + \beta_1x_1 + \beta_2x_2 + \beta_3 x_3 + \beta_4 x_4 + \varepsilon
-$$
-
-## Results
+[What methods/models are best]
+[What condition(s) does each method/model work well/poorly]
+[Conclusion about dataset]
 
@@
 ## Discussion
-Our model suggests that there is some weak evidence that with increasing temperature, the reproductive number R decreases.
-The effect of relative humidity on the reproductive number is much less conclusive.
-As we only used data from Italy in our model building, there is a limited range of weather information than we would like to have.
 

### 2020-04-18 PaperDraft.Rmd
epiestim and OLS part of methodology:

--- 2020-04-18 PaperDraft.Rmd
@@
 For each province, we take a 7 day rolling sum on the number of new cases each day, and feed the data into the parametric SI model, which is given distribution $w(\tau)$ From there, the model uses a likelihood based approach to estiamte R at each discrete timestep, computing the likelihood that a person at timestep $i$ is infected by a person infected at timestep $j$, and summing the likelihood for each $j$ across all its infectees to compute the $R$ value at timestep $j$.
+Thus, for each province, we produce a daily time series estimate for $R$ on each day.
 This method is outlined in detail in Wallinga and Teunis (2004)[8].
 
+For each province, day pair, we have $\bar{x} = (x_1, x_2, x_3, x_4), y$ where:
+
+- $x_1$: standardized GDP per capita
+- $x_2$: standardized population density
+- $x_3$: standardized air temperature
+- $x_4$: standardized relative humidity
+- $y$: logarithm of estimated R
+
+We then perform ordinary least squares regression to find $\bar{\beta}$:
+$$
+y = \beta_0 + \beta_1x_1 + \beta_2x_2 + \beta_3 x_3 + \beta_4 x_4 + \varepsilon
+$$
+
 ## Results
-
-[What methods/models are best]
-[What condition(s) does each method/model work well/poorly]
-[Conclusion about dataset]
-
-### Posterior Estimates
 
@@
 
-Fitting our Bayesian model, which extends the method of Thompson et. al, values
-of beta were estimated for temperature, relative humidity, Gross Domestic
-Product (GDP) and population density, respectively. The mean estimate of $\beta^T$ is
-
-$$ \beta^T = [-0.23198507 -0.05391226  0.26336295 -0.03586852] $$
-
-in the overall model for $R$:
-
+[TK: Beta Values]
+Fitting our Bayesian model, which extends the method of Thompson et.
+al, values of beta were estimated for temperature, relative humidity, Gross Domestic Product (GDP) and population density, respectively.
+Thus, our model estimates $\beta^T = [,,,,]$ in
 $$ R_i(t) = \exp\left( \theta(t) + \lambda_i + \mathbf{\beta}^T x_i(t) + \varepsilon_i(t) \right)$$
-
 Note that each feature's data is standardized to mean 0 and variance 1.
 
-The model indicates that higher temperatures are observed with lower $R$ values,
-and higher GDP is observed with higher $R$ values, while relative humidity and
-population density have close to no effect on $R$. Because of the confidence
-intervals, we are fairly confident of the directional correctness of this
-relationship for temperature and for GDP per capita.
+The model indicates that higher temperatures are observed with lower $R_0$ values, and higher GDP is observed with higher $R_0$ values, while relative humidity and population density have $R_0$ close to zero.
 
-### Estimates of $\theta$, transformed to climate-independent $R$ by date
+### Climate-Independent Basic R, by Date
 
@@
 
-$R$ is high in early phases, with an early spike on February 27, potentially due
-to increases in testing, which means the data (and hence the model) has
-underestimated the number of infectious cases compared to the ground truth.
-This result is aggregate across the entire nation and is net of any contribution
-to national weather changes.
+First, we observe that our computed value of $R_0$ in Italy matches what We observe that R_0 spikes in , with an early spike on February 27, potentially due to increased in testing, which captures cases before.
 
@@
 \includegraphics[width=.49\textwidth]{plots/plot9.png}
+\includegraphics[width=.49\textwidth]{plots/plot10.png}
 \end{center}
 
-These plots show the estimated $R$ values in each province computed through the
-samples from the MCMC. An important observation is that the confidence intervals
-for many provinces are very wide, because those provinces have relatively little
-data. However, one strength of our model is that no weighting on results from
-each province needs to be manually done. The data from larger provinces is
-automatically more important because the Poisson likelihood function returns
-smaller results as the rate parameter increases.
-
-We have also analyzed the $\lambda$ and $\varepsilon$ parameters of the model,
-but because these are not as important, see the Appendix for plots and analysis.
-
-In many cities, our estimated $R_0$ values are extremely high. This is because
-those cities reported a dramatic increase in cases, and our estimated $R_0$
-values are indeed consistent with the data. However, we will discuss in the
-interpretation section the limits of our approach, and why we believe this may
-not match the ground truth.
+Observe that in many cities, our estimated $R_0$ values are extremely high compared to the climate-independent plot shown earlier.
 
@@
 
-### Posterior of $\lambda$, location-dependent adjustment to $log(R)$
+### Posterior of Lambda, location-dependent adjustment to log(R)
 
@@
 
-$\lambda$ is an adjustment to the model based on the location. What we observe
-from the distribution of $\lambda$ is that some locations have much more
-certainty in how they deviate from the mean than others. This can be explained
-by the unequal distribution of cases across provinces; provinces with more cases
-lend themselves to more accurate estimation of $R$.
-
-### Epsilon, idiosyncratic adjustment to log(R)
-
+### Epsilon, Idiosyncratic adjustment to log(R)
 \begin{center}
@@
 \end{center}
-
-This histogram shows all samples of all entries in the $\varepsilon$ matrix.
-Generally, we prefer a lower average spread of $\varepsilon$, because it means
-that more of the variation in reproduction number can be predicted from the
-seasonal factors, location, and time alone. Unfortunately, we found the observed
-spread of $\varepsilon$ is quite high, indicating that there is substantial
-variance in reproductive numbers inferred from the data which cannot be
-explained by modeled features.
 

### 2020-04-18 PaperDraft.Rmd
Describe the data

--- 2020-04-18 PaperDraft.Rmd
@@
 
+### Data
+
+We sourced the following data:
+
+- Italian case data aggregated per province from the Italian Civil Protection
+  Department (DPC)
+- 2017 demographic data for Italian provinces from Eurostat, via the `restatapi` package
+- Weather data from NOAA's network of stations, via the `worldmet` package
+
+A total of $102$ provinces were included in the analysis. This is all provinces
+except those located in the region of Sardinia. The provinces in Sardinia were
+excluded because of a mismatch between the demographic data and the case data,
+due to a recent political reorganization in Sardinia.
+
+A total of $19$ dates were included in the analysis, starting from the first date
+of February 24, 2020 present in the data and concluding on March 13, 2020.
+Although more dates are available, we felt that the dramatic interventions in
+Italy after March 13 reduced the relevance of seasonal forcings on transmission.
+
+A total of $4$ demographic and climatic features were included in the analysis:
+air temperature, relative humidity, GDP per capita, and population density. The
+demographic features are included as control so that the inferred coefficients
+for weather are not due to spurious correlations with demographics, as Northern
+Italy is generally denser, richer, and cooler on average than Southern Italy.
+
 ## Results
@@
 
-$$ \beta^T = [-0.23198507 -0.05391226  0.26336295 -0.03586852] $$
+$$ \beta^T = \begin{bmatrix} -0.23198507 & -0.05391226 & 0.26336295 & -0.03586852 \end{bmatrix} $$
 

### 2020-04-18 PaperDraft.Rmd

@@
 For each province, we take a 7 day rolling sum on the number of new cases each day, and feed the data into the parametric SI model, which is given distribution $w(\tau)$ From there, the model uses a likelihood based approach to estiamte R at each discrete timestep, computing the likelihood that a person at timestep $i$ is infected by a person infected at timestep $j$, and summing the likelihood for each $j$ across all its infectees to compute the $R$ value at timestep $j$.
+Thus, for each province, we produce a daily time series estimate for $R$ on each day.
 This method is outlined in detail in Wallinga and Teunis (2004)[8].
+
+For each province, day pair, we have $\bar{x} = (x_1, x_2, x_3, x_4), y$ where:
+
+- $x_1$: standardized GDP per capita
+- $x_2$: standardized population density
+- $x_3$: standardized air temperature
+- $x_4$: standardized relative humidity
+- $y$: logarithm of estimated R
+
+We then perform ordinary least squares regression to find $\bar{\beta}$:
+$$
+y = \beta_0 + \beta_1x_1 + \beta_2x_2 + \beta_3 x_3 + \beta_4 x_4 + \varepsilon
+$$
 
@@
 ## Results
-
-[What methods/models are best]
-[What condition(s) does each method/model work well/poorly]
-[Conclusion about dataset]
 
@@
 ## Discussion
+Our model suggests that there is some weak evidence that with increasing temperature, the reproductive number R decreases.
+The effect of relative humidity on the reproductive number is much less conclusive.
+As we only used data from Italy in our model building, there is a limited range of weather information than we would like to have.
 

### 2020-04-18 PaperDraft.Rmd
small change

--- 2020-04-18 PaperDraft.Rmd
@@
 
-For each province, we take a 7 day rolling sum on the number of new cases each day, and feed the data into the parametric SI model, which is given distribution $w(\tau)$ From there, the model uses a likelihood based approach to estiamte R at each discrete timestep, computing the likelihood that a person at timestep $i$ is infected by a person infected at timestep $j$, and summing the likelihood for each $j$ across all its infectees to compute the $R$ value at timestep $j$.
+For each province, we take a 7 day rolling sum on the number of new cases each day, and feed the data into the parametric SI model, which is given distribution $w(\tau)$ based on the SI parameters.
+From there, the model uses a likelihood based approach to estiamte R at each discrete timestep, computing the likelihood that a person at timestep $i$ is infected by a person infected at timestep $j$, and summing the likelihood for each $j$ across all its infectees to compute the $R$ value at timestep $j$.
 Thus, for each province, we produce a daily time series estimate for $R$ on each day.
@@
 
-$$ \beta^T = [-0.23198507 -0.05391226  0.26336295 -0.03586852] $$
+$$ \beta^T = [-0.23198507, -0.05391226, 0.26336295, -0.03586852] $$
 
@@
 
-With our `cfc19i` library, it is
-
 Looking at the climate and infection data in American and Canadian cities as the situation develops further.

### 2020-04-18 PaperDraft.Rmd
update discussion and future research

--- 2020-04-18 PaperDraft.Rmd
@@
 ## Discussion
-Our model suggests that there is some weak evidence that with increasing temperature, the reproductive number R decreases.
+The posterior estimates of our model suggest that there is some weak evidence that with increasing temperature, the reproductive number R decreases.
 The effect of relative humidity on the reproductive number is much less conclusive.
 As we only used data from Italy in our model building, there is a limited range of weather information than we would like to have.
+With no access to line-list data to estimate the serial interval number ourselves, our estimated reproductive number R is limited by the accuracy of the serial interval number we used.
+The estimated R values tend to have wide confidence interval in the beginning dates due to the imported cases in the beginning and the lack of testing.
+Overall, our model does a good job estimating the reproductive number R in various provinces with high confidence level and provides valuable information for policy makers.
 
@@
 Look into statistical simulation models which consider intercity travel and geographic proximity.
+It would also be worthwhile to see the sensitivity of our model with different values of serial interval number.
 

### 2020-04-18 PaperDraft.Rmd
updated discussion

--- 2020-04-18 PaperDraft.Rmd
@@
 ## Discussion
-Our model suggests that there is some weak evidence that with increasing temperature, the reproductive number R decreases.
+The posterior estimates of our model suggest that there is some weak evidence that with increasing temperature, the reproductive number R decreases.
 The effect of relative humidity on the reproductive number is much less conclusive.
 As we only used data from Italy in our model building, there is a limited range of weather information than we would like to have.
+Our model does a good job estimating the reproductive number $R$ in various provinces with high confidence level and provides valuable information for policy makers.
+However, with no access to line-list data to estimate the serial interval number ourselves, our estimated reproductive number R is limited by the accuracy of the serial interval number we used.
+The estimated $R$ values tend to have wide confidence interval in the beginning dates due to the imported cases in the beginning and the lack of testing.
 
-### Directions for Future Research
+For furure research, we would like to look into the climate and infection data in different countries as the situation develops further.
+In particular, we would like to compare this with the results in Italy, and observe if the linear assumption between seasonal factors and $R$ holds true across different climate regions.
+If not, we would like to try fitting a nonlinear relationship between $R$ and seasonal factors.
 
-With our `cfc19i` library, it is
-
-Looking at the climate and infection data in American and Canadian cities as the situation develops further.
-In particular, compare this with the results in Italy, and observe if the linear assumption between seasonal factors and $R$ holds true across different climate regions.
-If not, try fitting while assuming a nonlinear relationship between $R$ and seasonal factors.
-
-Look into statistical simulation models which consider intercity travel and geographic proximity.
+We would also like to look into statistical simulation models which consider intercity travel and geographic proximity.
+As our estimated $R$ is dependent on the serial interval number, it would also be worthwhile to see the sensitivity of our model with different values of serial interval number as Flaxman has done in [6].
 

### 2020-04-18 PaperDraft.Rmd

--- 2020-04-18 PaperDraft.Rmd
@@
 
+```{r include=FALSE}
+res <- results()
+plot_beta_confidence(res)
+```
+
 Fitting our Bayesian model, which extends the method of Thompson et. al, values
@@
 
-$$ \beta^T = [-0.23198507, -0.05391226, 0.26336295, -0.03586852] $$
+$$ \beta^T = \begin{bmatrix} -0.232 & -0.054 & 0.263 & -0.036 \end{bmatrix} $$
 
@@
 relationship for temperature and for GDP per capita.
+
+In comparison, the EpiEstim and OLS model produced beta values of:
+
+$$ \beta^T = \begin{bmatrix} -0.051 & -0.045 & -0.081 & -0.051 \end{bmatrix} $$
+
+for temperature, relative humidity, Gross Domestic Product (GDP) and population density, respectively, with $$\beta_0 = 1.08802$$ in the overall model for $R$:
+
+$$
+R_i(t) = \exp\left(\beta_0 + \mathbf{\beta}^T x_i(t) + \varepsilon_i(t) \right)
+$$
 

### 2020-04-18 R/ThompsonModel.R
hotfix trash merge fix

@@ -1,21 +1,13 @@
 library(EpiEstim)
 library(dplyr)
 
-<<<<<<< HEAD
-=======
 #' Using Italian Provincial COVID-19 dataset augmented with climate and population statistics.
 #'
->>>>>>> 30e14e9ecd1b7ff762dcc262e488370aafdfae16
 #' Follows the framework of the Thompson et al. (2019) paper, while using OLS to predict R from climate similar to Wang et al.
 #' EpiEstim's parametric_si mode is actually based on work done by Wallinga and Teunis (2004).
 #' Primarily leverages the EpiEstim R package to perform MCMC and estimate R from time series.
 #' Requires some estimate of serial interval, which we simply use values from pervious papers for.
 #'
-<<<<<<< HEAD
-#' Using Italian Provincial COVID-19 dataset augmented with climate and population statistics.
-#' 
-=======
->>>>>>> 30e14e9ecd1b7ff762dcc262e488370aafdfae16
 #' @param filename The path to the augmented data file which we will be using.
 #' @return Fit object, result of linear regression on log(R) against standardized confounding variables.
 runThompson <- function(filename = "data/dpc-augmented.csv") {

### 2020-04-18 R/DataPipeline.R
non-ascii data never saved

@@ -27,27 +27,6 @@ fetch_file <- function(url, filename) {
   })
 }
 
-#' Fetches the latest COVID-19 cases data from European Center for Disease Control (ECDC).
-#'
-#' @returns None
-#'
-fetch_latest_ecdc <- function() {
-  url <- "https://opendata.ecdc.europa.eu/covid19/casedistribution/csv"
-  filename <- "data/ecdc-COVID-19-up-to-date.csv"
-  fetch_file(url, filename)
-
-  #' Quick cleanup on the column titles.
-  #'
-  d <- read.csv(filename, stringsAsFactors = FALSE)
-  d$t <- lubridate::decimal_date(as.Date(d$dateRep, format = "%d/%m/%Y"))
-  d <- d[order(d$'countriesAndTerritories', d$t, decreasing = FALSE), ]
-  names(d)[names(d) == "countriesAndTerritories"] <- "Countries.and.territories"
-  names(d)[names(d) == "deaths"] <- "Deaths"
-  names(d)[names(d) == "cases"] <- "Cases"
-  names(d)[names(d) == "dateRep"] <- "DateRep"
-  write.csv(d, filename)
-}
-
 #' Fetches latest data from John Hopkins University CSSE.
 #'
 #' @return None
@@ -67,13 +46,13 @@ fetch_latest_csse <- function() {
 }
 
 
-#' Fetch Italian provinces data from DPC.
+#' Fetch Italian provinces data from DPC. This file is never written as it contains non-ASCII characters.
 #'
 #' @return None
 #'
 fetch_latest_dpc <- function() {
   url <- "https://raw.githubusercontent.com/pcm-dpc/COVID-19/master/dati-province/dpc-covid19-ita-province.csv"
-  filename <- "data/dpc-covid19-ita-province.csv"
+  filename <- tempfile()
   fetch_file(url, filename)
 
   d <- read.csv(filename)
@@ -84,7 +63,7 @@ fetch_latest_dpc <- function() {
   names(d)[names(d) == "codice_provincia"] <- "province_code"
   names(d)[names(d) == "codice_regione"] <- "region_code"
   names(d)[names(d) == "totale_casi"] <- "total_cases"
-  write.csv(d, filename)
+  d
 }
 
 #' Obtain mapping of health ministry province names to eurostat codes
@@ -110,7 +89,7 @@ getprovincelist <- function() {
   )
[…]

### 2020-04-18 PaperDraft.Rmd

--- 2020-04-18 PaperDraft.Rmd
@@
 
-$$ \beta^T = \begin{bmatrix} -0.23198507 & -0.05391226 & 0.26336295 & -0.03586852 \end{bmatrix} $$
+$$ \beta^T = \begin{bmatrix} -0.232 & -0.054 & 0.263 & -0.036 \end{bmatrix} $$
 
@@
 ## Discussion
+
 The posterior estimates of our model suggest that there is some weak evidence that with increasing temperature, the reproductive number R decreases.

### 2020-04-18 PaperDraft.Rmd
some formatting in reprot

--- 2020-04-18 PaperDraft.Rmd
@@
 Amidst the current COVID-19 epidemic, the prediction of its future spread and the length of the epidemic becomes increasingly significant.
-We investigate the effects of temperature and humidity on the reproductive rate R of COVID-19, as warmer climates in the summer have historically resulted in diminished spread of influenza.
-Early epidemiology research in COVID-19 have used simple linear models such as ordinary least squares regression to estimate the relationship between R and climate.
+We investigate the effects of temperature and humidity on the reproductive rate $R$ of COVID-19, as warmer climates in the summer have historically resulted in diminished spread of influenza.
+Early epidemiology research in COVID-19 have used simple linear models such as ordinary least squares regression to estimate the relationship between $R$ and climate.
 We estimate the reproductive rate of the virus using a novel Bayesian inference model with MCMC sampling to daily case data of various climate regions.
-While results are largely inconclusive, they provide us with a weak posterior update that R decreases with temperature and relative humidity.
+While results are largely inconclusive, they provide us with a weak posterior update that $R$ decreases with temperature and relative humidity.
 
@@
 The estimation of the latter quantity is important for policy, because infectious disease outbreaks will switch from an exponential growth regime to an exponential decay regime when the reproductive number falls below 1.
-The goal of mitigation or suppression policies is therefore to effect change on the reproductive number R.
+The goal of mitigation or suppression policies is therefore to effect change on the reproductive number $R$.
 Many infectious diseases are believed to be seasonal, including most respiratory viruses, like the SARS-CoV-2 virus implicated in the 2019–2020 coronavirus pandemic.
 Wang et al.
-(2020) [3] have analysed how the weather of locations within China correlates with local point estimates of the reproductive number R.
+(2020) [3] have analysed how the weather of locations within China correlates with local point estimates of the reproductive number $R$.
 However, because their analysis is frequentist in nature, their estimates of the seasonal forcing do not come with any estimate of uncertainty.
@@
 (2020) [1], which estimated the mean serial interval to be 4.7 and the standard deviation to be 2.4.
-We use the model detailed in Wallinga and Teunis (2004) [8] for estimating R from bayesian inference on a fixed serial interval distribution and time series data, which can be easily implemented using the EpiEstim package produced by Cori et al.
+We use the model detailed in Wallinga and Teunis (2004) [8] for estimating $R$ from bayesian inference on a fixed serial interval distribution and time series data, which can be easily implemented using the EpiEstim package produced by Cori et al.
 (2013) [7].
 
-For each province, we take a 7 day rolling sum on the number of new cases each day, and feed the data into the parametric SI model, which is given distribution $w(\tau)$ From there, the model uses a likelihood based approach to estiamte R at each discrete timestep, computing the likelihood that a person at timestep $i$ is infected by a person infected at timestep $j$, and summing the likelihood for each $j$ across all its infectees to compute the $R$ value at timestep $j$.
+For each province, we take a 7 day rolling sum on the number of new cases each day, and feed the data into the parametric SI model, which is given distribution $w(\tau)$ From there, the model uses a likelihood based approach to estiamte $R$ at each discrete timestep, computing the likelihood that a person at timestep $i$ is infected by a person infected at timestep $j$, and summing the likelihood for each $j$ across all its infectees to compute the $R$ value at timestep $j$.
 Thus, for each province, we produce a daily time series estimate for $R$ on each day.
@@
 - $x_4$: standardized relative humidity
-- $y$: logarithm of estimated R
+- $y$: logarithm of estimated $R$
 
@@
 ## Discussion
-The posterior estimates of our model suggest that there is some weak evidence that with increasing temperature, the reproductive number R decreases.
+The posterior estimates of our model suggest that there is some weak evidence that with increasing temperature, the reproductive number $R$ decreases.
 The effect of relative humidity on the reproductive number is much less conclusive.
@@
 Our model does a good job estimating the reproductive number $R$ in various provinces with high confidence level and provides valuable information for policy makers.
-However, with no access to line-list data to estimate the serial interval number ourselves, our estimated reproductive number R is limited by the accuracy of the serial interval number we used.
+However, with no access to line-list data to estimate the serial interval number ourselves, our estimated reproductive number $R$ is limited by the accuracy of the serial interval number we used.
 The estimated $R$ values tend to have wide confidence interval in the beginning dates due to the imported cases in the beginning and the lack of testing.
@@
 We would also like to look into statistical simulation models which consider intercity travel and geographic proximity.
-As our estimated $R$ is dependent on the serial interval number, it would also be worthwhile to see the sensitivity of our model with different values of serial interval number as Flaxman has done in [6].
+As our estimated $R$ is dependent on the serial interval number, it would also be worthwhile to see the sensitivity of our model with different values of serial interval number as did Flaxman et al.
+(2020)[6].
 

### 2020-04-18 R/AnalysisDetails.R
more to ignore

@@ -2,18 +2,18 @@
 #' Must be called after `collectData()` and `runMCMC()` and `runThompson()`.
 #' @return The saved results of the analysis.
 results <- function() {
-  if (!file.exists("data/dpc-augmented.csv")) {
+  if (!file.exists("inst/extdata/dpc-augmented.csv")) {
     throw.error("Could not find augmented DPC data. Did you `collectData()`?")
-  } else if (!file.exists("data/cf_fit.rds") | !file.exists("data/cf_data.rds") | !file.exists("data/cf_dates.rds")) {
+  } else if (!file.exists("inst/extdata/cf_fit.rds") | !file.exists("inst/extdata/cf_data.rds") | !file.exists("inst/extdata/cf_dates.rds")) {
     throw.error("Could not find MCMC fit results. Did you `runMCMC()`?")
-  } else if (!file.exists("data/thompson_fit.rds")) {
+  } else if (!file.exists("inst/extdata/thompson_fit.rds")) {
     throw.error("Could not find Thompson fit results. Did you `runThompson()`?")
   }
-  dpc <- read.csv("data/dpc-augmented.csv")
-  load("data/cf_fit.rds")
-  load("data/cf_data.rds")
-  load("data/cf_dates.rds")
-  load("data/thompson_fit.rds")
+  dpc <- read.csv("inst/extdata/dpc-augmented.csv")
+  load("inst/extdata/cf_fit.rds")
+  load("inst/extdata/cf_data.rds")
+  load("inst/extdata/cf_dates.rds")
+  load("inst/extdata/thompson_fit.rds")
   list(
     descr = c("temp", "RH", "GDP per capita", "density"),
     dpc = dpc,

### 2020-04-18 R/AnalysisDetails.R
more changes

@@ -1,6 +1,7 @@
 #' Return a list containing results of the analysis, sufficient for generating plots.
 #' Must be called after `collectData()` and `runMCMC()` and `runThompson()`.
 #' @return The saved results of the analysis.
+#' @export
 results <- function() {
   if (!file.exists("inst/extdata/dpc-augmented.csv")) {
     throw.error("Could not find augmented DPC data. Did you `collectData()`?")
@@ -25,4 +26,3 @@ results <- function() {
     thompsonfit = fit
   )
 }
-

### 2020-04-18 R/ClimateForcingModel.R

@@ -26,6 +26,7 @@ getSIdist <- function(n, mean = 4.7, sd = 2.9) {
 
 #' Compile the stan model, run MCMC, and save results.
 #' @param cutoff The cutoff date (default: Mar 14).
+#' @export
 runMCMC <- function(cutoff = "2020-03-14") {
   df <- read.csv("data/dpc-augmented.csv", stringsAsFactors = FALSE)
 

### 2020-04-18 R/DataDocumentation.R

@@ -1,17 +1,7 @@
-#' demodata
-#'
-#' Demographic data from Eurostat, including GDP per capita, population, and
-#' population density.
-#' @docType data
-#' @name demodata
-NULL
-
-#' dpc-augmented
-#'
-#' Processed data including weather, demographics, and also the DPC Italy
-#' disease incidence data.
-#' @docType data
-#' @name dpc-augmented
+#' @importFrom stats density dnorm filter lm na.omit quantile
+#' @importFrom utils download.file read.csv write.csv
+#' @importFrom dplyr %>% coalesce filter first full_join group_by group_map
+#'   group_modify left_join mutate rename ungroup
 NULL
 
 #' dpc-covid19-ita-province
@@ -19,14 +9,6 @@ NULL
 #' Raw data per Italian province published by the Italy DPC.
 #' @docType data
 #' @name dpc-covid19-ita-province
-NULL
-
-#' ecdc-COVID-19-up-to-date
-#'
-#' Raw data for various countries published by the ECDC. Not used in the
-#' analysis by default.
-#' @docType data
-#' @name ecdc-COVID-19-up-to-date
 NULL
 
 #' time_series_covid19_confirmed_global

### 2020-04-18 R/DataPipeline.R

@@ -297,6 +297,7 @@ transform_total_cases <- function(dpc) {
 #' @param rewriteall Attempt to refresh all files even if already exist. If
 #'   TRUE, this function may need to be called multiple times due to `worldmet`
 #'   issues.
+#' @export
 collectData <- function(rewriteall = FALSE) {
   fetch_latest_csse()
 

### 2020-04-18 R/Plots.R

@@ -33,6 +33,7 @@ removeyear <- function(range) {
 #' @param res Output of `results()` containing results of the analysis.
 #' @return A plot of location of standardized beta parameters, with 50%
 #'   confidence interval highlighted.
+#' @export
 plot_beta_confidence <- function(res) {
   betadescr <- paste(res$descr, sep = ", ", collapse = ", ")
   (
@@ -53,6 +54,7 @@ plot_beta_confidence <- function(res) {
 #' @param end Last location to display.
 #' @return A plot of location of standardized beta parameters, with 50%
 #'   confidence interval highlighted.
+#' @export
 plot_lambda_confidence <- function(res, start, end) {
   (
     mcmc_intervals(
@@ -81,6 +83,7 @@ plot_lambda_confidence <- function(res, start, end) {
 #'
 #' @param res Samples of posterior distribution of parameters.
 #' @return A ggplot2 time series of estimated independent R with 80% confidence.
+#' @export
 plot_independentR <- function(res) {
   thetadf <- data.frame(res$params$theta)
   colnames(thetadf) <- removeyear(res$dates)
@@ -109,6 +112,7 @@ plot_independentR <- function(res) {
 #' @param res Samples of posterior distribution of parameters.
 #' @param loc Index of the location to examine.
 #' @return A ggplot2 time series of estimated R with 80% confidence.
+#' @export
 plot_locationR <- function(res, loc) {
   thetadf <- data.frame(res$params$theta +
                           res$params$lambda[loc] +
@@ -138,6 +142,7 @@ plot_locationR <- function(res, loc) {
 #'
 #' @param res Samples of posterior distribution of parameters.
 #' @return A histogram of all samples of epsilon.
+#' @export
 plot_epsilon_dispersion <- function(res) {
   df <- data.frame(adjustment = as.vector(res$epsilon))
   ggplot(df, aes(adjustment)) +
@@ -147,18 +152,3 @@ plot_epsilon_dispersion <- function(res) {
                   xlab = "difference in log(R)")
 }
 
-# TODO: move these to the Rmarkdown file
-generate_plots <- function() {
-  res <- results()
-  plot_beta_confidence(res)
-  plot_lambda_confidence(res, 1, 34)
-  plot_lambda_confidence(res, 35, 68)
-  plot_lambda_confidence(res, 69, 102)
-  plot_independentR(res)
-  plot_locationR(res, 13)
[…]

### 2020-04-18 R/ThompsonModel.R

@@ -10,6 +10,7 @@ library(dplyr)
 #'
 #' @param filename The path to the augmented data file which we will be using.
 #' @return Fit object, result of linear regression on log(R) against standardized confounding variables.
+#' @export
 runThompson <- function(filename = "inst/extdata/dpc-augmented.csv") {
   dpc_df <- read.csv(filename)
   dpc_df$date <- as.Date(dpc_df$date, format = "%Y-%m-%d")

### 2020-04-18 PaperDraft.Rmd
Update PaperDraft.Rmd

@@
 
+For each province, we take a 7 day rolling sum on the number of new cases each day, and feed the data into the parametric SI model, which is given distribution $w(\tau)$ based on the SI parameters.
+From there, the model uses a likelihood based approach to estiamte R at each discrete timestep, computing the likelihood that a person at timestep $i$ is infected by a person infected at timestep $j$, and summing the likelihood for each $j$ across all its infectees to compute the $R$ value at timestep $j$.
+Thus, for each province, we produce a daily time series estimate for $R$ on each day.
+This method is outlined in detail in Wallinga and Teunis (2004)[8].
+
 For each province, day pair, we have $\bar{x} = (x_1, x_2, x_3, x_4), y$ where:
@@
 
+```{r include=FALSE}
+```
+
 Fitting our Bayesian model, which extends the method of Thompson et. al, values
@@
 relationship for temperature and for GDP per capita.
+
+In comparison, the EpiEstim and OLS model produced beta values of:
+
+$$ \beta^T = \begin{bmatrix} -0.051 & -0.045 & -0.081 & -0.051 \end{bmatrix} $$
+
+for temperature, relative humidity, Gross Domestic Product (GDP) and population density, respectively, with $$\beta_0 = 1.08802$$ in the overall model for $R$:
+
+$$
+R_i(t) = \exp\left(\beta_0 + \mathbf{\beta}^T x_i(t) + \varepsilon_i(t) \right)
+$$
+
+We note that the coefficient for GDP per capita has flipped signs from the Bayesian framework estimate, with the other coefficient remaining the same sign.
+The p-values for GDP per capita and density are 0.00042 and 0.019 respectively, which are both significant.
+However, the p-values for air temperature and relative humidity are 0.42 and 0.65 respectively, which are much less significant.
+The large discrepancy in values between coefficients between the two models is surprising, considering they are both based on a linear relationship between $\log(R)$ and the confounding varaibles.
+Further investigation would be required in the future to determine the cause.
 
@@
 ## Discussion
+
 The posterior estimates of our model suggest that there is some weak evidence that with increasing temperature, the reproductive number $R$ decreases.
@@
 
-For furure research, we would like to look into the climate and infection data in different countries as the situation develops further.
+For future research, we would like to look into the climate and infection data in different countries as the situation develops further.
 In particular, we would like to compare this with the results in Italy, and observe if the linear assumption between seasonal factors and $R$ holds true across different climate regions.

### 2020-04-18 R/ClimateForcingModel.R
fixup: data move

@@ -28,7 +28,7 @@ getSIdist <- function(n, mean = 4.7, sd = 2.9) {
 #' @param cutoff The cutoff date (default: Mar 14).
 #' @export
 runMCMC <- function(cutoff = "2020-03-14") {
-  df <- read.csv("data/dpc-augmented.csv", stringsAsFactors = FALSE)
+  df <- read.csv("inst/extdata/dpc-augmented.csv", stringsAsFactors = FALSE)
 
   # Stop the analysis on Mar 13
   df <- filter(df, date < cutoff)

### 2020-04-18 R/DataDocumentation.R

@@ -1,7 +1,16 @@
-#' @importFrom stats density dnorm filter lm na.omit quantile
+#' @importFrom stats density dnorm lm na.omit quantile
 #' @importFrom utils download.file read.csv write.csv
 #' @importFrom dplyr %>% coalesce filter first full_join group_by group_map
-#'   group_modify left_join mutate rename ungroup
+#'   group_modify left_join mutate rename select summarize ungroup
+#' @importFrom memoise has_cache
+#' @importFrom rstan sampling
+#' @importFrom abind abind
+#' @importFrom bayesplot mcmc_areas mcmc_intervals
+#' @importFrom ggplot2 element_blank element_text geom_histogram geom_smooth ggplot xlim ylab
+#' @importFrom worldmet getMeta
+#' @importFrom restatapi get_eurostat_data get_eurostat_dsd
+#' @importFrom EpiEstim estimate_R
+#' @importFrom R.utils withTimeout
 NULL
 
 #' dpc-covid19-ita-province

### 2020-04-18 R/DataDocumentation.R
fix NAMESPACE again

@@ -6,11 +6,12 @@
 #' @importFrom rstan sampling
 #' @importFrom abind abind
 #' @importFrom bayesplot mcmc_areas mcmc_intervals
-#' @importFrom ggplot2 element_blank element_text geom_histogram geom_smooth ggplot xlim ylab
+#' @importFrom ggplot2 aes element_blank element_text geom_histogram geom_smooth ggplot xlim ylab
 #' @importFrom worldmet getMeta
 #' @importFrom restatapi get_eurostat_data get_eurostat_dsd
-#' @importFrom EpiEstim estimate_R
+#' @importFrom EpiEstim estimate_R make_config
 #' @importFrom R.utils withTimeout
+#' @importFrom tidyr pivot_longer
 NULL
 
 #' dpc-covid19-ita-province
@@ -43,3 +44,8 @@ NULL
 #' @docType data
 #' @name time_series_covid19_recovered_global
 NULL
+
+# dplyr "globals" and cf_data, cf_fit which come from `load`, throw.error from R.oo
+utils::globalVariables(c("RH", "adjustment", "age", "air_temp", "cf_data", "cf_fit", "code", "concept",
+                         "dates", "fit", "gdppercapita", "lat", "logR", "province", "region", "sex",
+                         "theme", "throw.error", "total_cases", "unit", "values", "values.x", "values.y"))

### 2020-04-18 R/DataDocumentation.R
another global

@@ -48,4 +48,5 @@ NULL
 # dplyr "globals" and cf_data, cf_fit which come from `load`, throw.error from R.oo
 utils::globalVariables(c("RH", "adjustment", "age", "air_temp", "cf_data", "cf_fit", "code", "concept",
                          "dates", "fit", "gdppercapita", "lat", "logR", "province", "region", "sex",
-                         "theme", "throw.error", "total_cases", "unit", "values", "values.x", "values.y"))
+                         "station", "theme", "throw.error", "total_cases", "unit", "values", "values.x",
+                         "values.y"))

### 2020-04-18 PaperDraft.Rmd

@@
 
-Amidst the current COVID-19 epidemic, the prediction of its future spread and the length of the epidemic becomes increasingly significant.
+Amidst the current COVID-19 epidemic, the prediction of its future spread and the length of the epidemic become increasingly significant.
 We investigate the effects of temperature and humidity on the reproductive rate $R$ of COVID-19, as warmer climates in the summer have historically resulted in diminished spread of influenza.
@@
 The estimation of the latter quantity is important for policy, because infectious disease outbreaks will switch from an exponential growth regime to an exponential decay regime when the reproductive number falls below 1.
-The goal of mitigation or suppression policies is therefore to effect change on the reproductive number $R$.
-Many infectious diseases are believed to be seasonal, including most respiratory viruses, like the SARS-CoV-2 virus implicated in the 2019–2020 coronavirus pandemic.
+The goal of mitigation or suppression policies is therefore to reduce the reproductive number $R$.
+Many infectious diseases are seasonal, including most respiratory viruses, like the SARS-CoV-2 virus implicated in the 2019–2020 coronavirus pandemic.
 Wang et al.
@@
 $R$ through MCMC. This involves both substantial deviations from the second part
-of the Thompson et al. model and also less less substantial adjustments for the
+of the Thompson et al. model and also less substantial adjustments for the
 nature of our data. Our method does not extend the first part of the Thompson et

### 2020-04-18 PaperDraft.Rmd
Remove unnecessary Thompson fit results

@@
 consensus of international estimates for $R_0$, which vary widely but tend to
-cluster between $2$ or $3$ [TK: Citation needed?]. Because of the great
+cluster between $2$ or $3$. Because of the great
 uncertainty, there is significant variance in this assumption, so the prior is
@@
 ```{r include=FALSE}
+require(cfc19i)
+
+res <- results()
 ```

### 2020-04-18 R/AnalysisDetails.R

@@ -7,14 +7,11 @@ results <- function() {
     throw.error("Could not find augmented DPC data. Did you `collectData()`?")
   } else if (!file.exists("inst/extdata/cf_fit.rds") | !file.exists("inst/extdata/cf_data.rds") | !file.exists("inst/extdata/cf_dates.rds")) {
     throw.error("Could not find MCMC fit results. Did you `runMCMC()`?")
-  } else if (!file.exists("inst/extdata/thompson_fit.rds")) {
-    throw.error("Could not find Thompson fit results. Did you `runThompson()`?")
   }
   dpc <- read.csv("inst/extdata/dpc-augmented.csv")
   load("inst/extdata/cf_fit.rds")
   load("inst/extdata/cf_data.rds")
   load("inst/extdata/cf_dates.rds")
-  load("inst/extdata/thompson_fit.rds")
   list(
     descr = c("temp", "RH", "GDP per capita", "density"),
     dpc = dpc,
@@ -22,7 +19,6 @@ results <- function() {
     standata = cf_data,
     params = rstan::extract(cf_fit),
     dates = dates,
-    locations = colnames(cf_data$i),
-    thompsonfit = fit
+    locations = colnames(cf_data$i)
   )
 }

### 2020-04-18 R/AnalysisDetails.R
Remove errors

@@ -3,11 +3,6 @@
 #' @return The saved results of the analysis.
 #' @export
 results <- function() {
-  if (!file.exists("inst/extdata/dpc-augmented.csv")) {
-    throw.error("Could not find augmented DPC data. Did you `collectData()`?")
-  } else if (!file.exists("inst/extdata/cf_fit.rds") | !file.exists("inst/extdata/cf_data.rds") | !file.exists("inst/extdata/cf_dates.rds")) {
-    throw.error("Could not find MCMC fit results. Did you `runMCMC()`?")
-  }
   dpc <- read.csv("inst/extdata/dpc-augmented.csv")
   load("inst/extdata/cf_fit.rds")
   load("inst/extdata/cf_data.rds")

### 2020-04-18 R/DataDocumentation.R
remove throw.error

@@ -45,8 +45,7 @@ NULL
 #' @name time_series_covid19_recovered_global
 NULL
 
-# dplyr "globals" and cf_data, cf_fit which come from `load`, throw.error from R.oo
+# dplyr "globals" and cf_data, cf_fit which come from `load`
 utils::globalVariables(c("RH", "adjustment", "age", "air_temp", "cf_data", "cf_fit", "code", "concept",
                          "dates", "fit", "gdppercapita", "lat", "logR", "province", "region", "sex",
-                         "station", "theme", "throw.error", "total_cases", "unit", "values", "values.x",
-                         "values.y"))
+                         "station", "theme", "total_cases", "unit", "values", "values.x", "values.y"))

### 2020-04-18 R/ThompsonModel.R
thompson

@@ -9,7 +9,6 @@ library(dplyr)
 #'
 #' @param filename The path to the augmented data file which we will be using.
 #' @return Fit object, result of linear regression on log(R) against standardized confounding variables.
-#' @export
 runThompson <- function(filename = "inst/extdata/dpc-augmented.csv") {
   dpc_df <- read.csv(filename)
   dpc_df$date <- as.Date(dpc_df$date, format = "%Y-%m-%d")

### 2020-04-18 PaperDraft.Rmd
Remove attempt at inline plots
add captions appendix

@@
 
-```{r include=FALSE}
-require(cfc19i)
-
-res <- results()
-```
-
 Fitting our Bayesian model, which extends the method of Thompson et. al, values
@@
 
-\begin{center}
+\begin{figure}
 \includegraphics[width=.49\textwidth]{plots/plot2.png}
@@
 \includegraphics[width=.49\textwidth]{plots/plot4.png}
-\end{center}
+\caption{Posterior location of $\lambda$, split into three plots for each
+location. $\lambda$ is an adjustment to the model based on the location.}
+\end{figure}
 
-$\lambda$ is an adjustment to the model based on the location. What we observe
-from the distribution of $\lambda$ is that some locations have much more
-certainty in how they deviate from the mean than others. This can be explained
-by the unequal distribution of cases across provinces; provinces with more cases
-lend themselves to more accurate estimation of $R$.
+What we observe from the distribution of $\lambda$ is that some locations have
+much more certainty in how they deviate from the mean than others. This can be
+explained by the unequal distribution of cases across provinces; provinces with
+more cases lend themselves to more accurate estimation of $R$.
 
@@
 
-\begin{center}
+\begin{figure}
 \includegraphics[width=.49\textwidth]{plots/plot11.png}
-\end{center}
+\caption{Histogram of all samples of all entires in the $\varepsilon$ matrix.}
+\end{figure}
 
-This histogram shows all samples of all entries in the $\varepsilon$ matrix.
 Generally, we prefer a lower average spread of $\varepsilon$, because it means

### 2020-04-18 PaperDraft.Rmd
figures

--- 2020-04-18 PaperDraft.Rmd
@@
 author: "Fengyang Wang, Lu Lu, Ziyao Tian, Yi Xiao Zhang"
+organization: "STAT 440 Final Project"
 header-includes:
@@
 output:
-  pdf_document: default
+  pdf_document:
+    number_sections: true
   html_notebook: default
-  rmarkdown::pdf_document:
-    fig_caption: yes        
-    includes:  
-      in_header: preamble.tex
 ---
 
-## Abstract
+\textbf{Abstract}
 
@@
 
-## Introduction
+# Introduction
 
@@
 
-## Methodology
+# Methodology
 
@@
 
-### Incidence likelihood of the Thompson et al. model
+## Incidence likelihood of the Thompson et al. model
 
@@
 
-### Extension of Thompson et al. for $R$ subject to forcings
+## Extension of Thompson et al. for $R$ subject to forcings
 
@@
 
-### Using OLS and EpiEstim to check results
+## Using OLS and EpiEstim to check results
 
@@
 
-### Data
+## Data
 
@@
 
-## Results
+# Results
 
-### Posterior Estimates
+## Posterior Estimates
 
@@
 
-\begin{center}
+\begin{figure}
+\centering
 \includegraphics[width=0.5\textwidth]{plots/plot1.png}
-\end{center}
-
-```{r include=FALSE}
-```
+\caption{Posterier distributions of beta}
+\end{figure}
 
@@
 
-### Estimates of $\theta$, transformed to climate-independent $R$ by date
+## Estimates of $\theta$, transformed to climate-independent $R$ by date
 
-\begin{center}
+\begin{figure}
+\centering
 \includegraphics[width=.49\textwidth]{plots/plot5.png}
-\end{center}
+\caption{Time series estimate of R with 0.8 confidence bands}
+\end{figure}
 
@@
 
-### Estimated R in Various Provinces
+## Estimated R in Various Provinces
 
-\begin{center}
+\begin{figure}
+\centering
 \includegraphics[width=.49\textwidth]{plots/plot6.png}
@@
 \includegraphics[width=.49\textwidth]{plots/plot9.png}
-\end{center}
+\caption{Time series estimates of R for Bergano, Milan, Rome, Venice}
+\end{figure}
 
@@
 
-### Interpretation
+## Interpretation
 
@@
 
-## Discussion
+# Discussion
 
@@
 (2020)[6].
+\newpage
+# Appendix
 
-## Appendix
+## Posterior of $\lambda$, location-dependent adjustment to $log(R)$
 
-### Posterior of $\lambda$, location-dependent adjustment to $log(R)$
-
-\begin{center}
+\begin{figure}
 \includegraphics[width=.49\textwidth]{plots/plot2.png}
@@
 \includegraphics[width=.49\textwidth]{plots/plot4.png}
-\end{center}
+\end{figure}
 
@@
 
-### Epsilon, idiosyncratic adjustment to log(R)
+## Epsilon, idiosyncratic adjustment to log(R)
 
@@
 
-### References
+## References
 

### 2020-04-18 PaperDraft.Rmd

@@
 (2020)[6].
+
 \newpage
+
 # Appendix
@@
 \begin{figure}
+\centering
 \includegraphics[width=.49\textwidth]{plots/plot2.png}
@@
 \includegraphics[width=.49\textwidth]{plots/plot4.png}
+\caption{Posterior location of $\lambda$, split into three plots for each
+location. $\lambda$ is an adjustment to the model based on the location.}
 \end{figure}
 
-$\lambda$ is an adjustment to the model based on the location. What we observe
-from the distribution of $\lambda$ is that some locations have much more
-certainty in how they deviate from the mean than others. This can be explained
-by the unequal distribution of cases across provinces; provinces with more cases
-lend themselves to more accurate estimation of $R$.
+What we observe from the distribution of $\lambda$ is that some locations have
+much more certainty in how they deviate from the mean than others. This can be
+explained by the unequal distribution of cases across provinces; provinces with
+more cases lend themselves to more accurate estimation of $R$.
 
@@
 
-\begin{center}
+\begin{figure}
+\centering
 \includegraphics[width=.49\textwidth]{plots/plot11.png}
-\end{center}
+\caption{Histogram of all samples of all entires in the $\varepsilon$ matrix.}
+\end{figure}
 
-This histogram shows all samples of all entries in the $\varepsilon$ matrix.
 Generally, we prefer a lower average spread of $\varepsilon$, because it means
